---
title: "1_Data_Manipulation"
author: "Alex Dhond"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to load in the PREDICTS data and filter it down so I can get data about the species and sites for Neotropical ants, dung beetles, and the combination of both. Some of the code was provided by Adriana De Palma for use in processing the PREDICTS data, which I have adapted here.

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(taxize) #for linking the NCBI taxonomy to check if the dung beetles I have are dung beetles
```

### Load in the most updated version of the PREDICTS database
```{r}
diversity <- readRDS("UPDATED.rds")
```

### See which studies are missing sampling efforts
```{r}
studies_check <- diversity %>%
  # filter the rows where Sampling efforts are NA
  filter(is.na(Sampling_effort)) %>%
  # keep only unique studies
  distinct(SS) %>%
  # pull the vector
  pull(SS)

diversity %>%
  # filter the rows where studies are those that had missing sampling efforts (above)
  filter(SS %in% studies_check) %>%
  # drop missing levels
  droplevels() %>%
  # pull out the sampling efforts of these studies
  pull(Sampling_effort)%>%
  # summarise to check that ALL the data are NAs
  summary()
```

### Replace missing sampling efforts with 1
```{r}
diversity <- diversity %>%
  
  # replace missing sampling effort values with 1
  mutate(x = replace_na(Sampling_effort, 1)) %>%
  
  # group by Study
  group_by(SS) %>%
  
  # check how many sampling efforts there are in each study
  mutate(n_sample_effort = n_distinct(Sampling_effort),
         # get the maximum sampling effort for the studies
         
         max_sample_effort = max(Sampling_effort)
         ) %>%
  
  ungroup() %>%
  
  # if the study has more than one sampling effort, correct the abundance

    # so if there's only one sampling effort, then leave the dividing effort as empty because we don't want to change the abundances when we do the divisions. Otherwise, we give it the maximum sampling effort.
  mutate(DividingEffort = ifelse(n_sample_effort == 1, NA, max_sample_effort)) %>%
  
  # if the diversity metric isn't sensitive to the effort, then we'll change the value to NA too (so we won't end up changing the measurement), otherwise leave it as it is
  mutate(DividingEffort = ifelse(Diversity_metric_is_effort_sensitive == FALSE, NA, DividingEffort)) %>%
  
  # now let's create the effort corrected measurement by dividing the abundances by the sampling efforts
  # where the dividing effort isn't NA (i.e. when it is necessary)
  mutate(Corrected_sampling_effort = ifelse(is.na(DividingEffort), 
                                            Sampling_effort,
                                            Sampling_effort / DividingEffort),
         Effort_corrected_measurement = ifelse(is.na(DividingEffort),
                                               Measurement,
                                               Measurement * Corrected_sampling_effort))
  
# summarise the corrected sampling efforts
summary(diversity$Corrected_sampling_effort)
```

### Test that. Here, there are multiple sampling efforts in this source
```{r}
diversity %>% 
  # filter out the test study
  filter(SS == "AD1_2001__Liow 1") %>% 
  # select out the columns to check
  dplyr::select(Measurement, Effort_corrected_measurement, Sampling_effort, Corrected_sampling_effort)
```


### Here, the effort corrected measurement and the measurement should be the same, since there is only one sampling effort
```{r}
diversity %>% 
  filter(SS == "AD1_2004__Darvill 1") %>% 
  dplyr::select(Measurement, Effort_corrected_measurement,
                Sampling_effort, Corrected_sampling_effort)
```

### Test a study of species richness to ensure it works
```{r}
diversity %>% 
  # filter out the test study
  filter(SS == "AD1_2011__Holzschuh 2") %>% 
  # select out the columns to check
  dplyr::select(Measurement, Effort_corrected_measurement,
                Sampling_effort, Corrected_sampling_effort)
```

### Merge identical sites together
Merge any sites that are the same land-use type and have identical coordinates, sampling start and end dates
```{r}
diversity <- diversity %>%
  
  # group by aspects of the sites that should be identical if we need to merge the abundances
  # I only want to merge abundances if they are within the same study and block
  # as I'm assuming that even if the locations and sampling times are the same, if the blocks or studies are different, then there is some good reason for this.
  group_by(Source_ID, Study_number, Study_name, Block,
           #diversity metric type
           Diversity_metric, Diversity_metric_type, Diversity_metric_unit,
           Diversity_metric_is_effort_sensitive,
           
           #details of the sites
           Predominant_habitat, Use_intensity, Years_since_fragmentation_or_conversion,
           
           #details of the sampling method
           Sampling_method, Sampling_effort_unit,
           
           #species identity
           Study_common_taxon, Rank_of_study_common_taxon,
           Taxon_number, Taxon_name_entered,
           Indication, Parsed_name,
           Best_guess_binomial, COL_ID, Taxon, Name_status,
           Rank, Kingdom, Phylum, Class, Order, Family, Genus, Species,
           Higher_taxon,
           
           #site location
           Longitude, Latitude,
           
           #sampling time
           Sample_start_earliest, Sample_end_latest, Sample_date_resolution) %>%
  
  # if the diversity metric is occurrence:
  #   if it is present at all, give it a 1, if it is always absent, give it a 0,
  # otherwise (if the metric is either abundance or species richness):
  #   calculate the weighted abundance/richness for each taxonomic group, weighted by sampling effort
  
  mutate(merged_diversity = 
           ifelse(Diversity_metric_type == "Occurrence",
                  # if any of the occurrence values are 1, `any` will return TRUE. If you sum a logical, TRUE becomes 1 and FALSE becomes 0
                  sum(any(Effort_corrected_measurement > 0)),
                  
                  # note that since we've already corrected the sampling effort, this is essentially a mean rather than a weighted mean for abundance measurements. It's a weighted mean for species richness though where sampling efforts vary.
                  stats::weighted.mean(x = Effort_corrected_measurement,
                                       w = Corrected_sampling_effort))
         )

# pull out the grouping data (so we can double check how many records we're merging for each)
group_dat <- diversity %>% 
  group_data() %>%
  mutate(nvals_merged = lengths(.rows),
         merge_ID = row_number())

# ungroup the diversity data for future use
diversity <- ungroup(diversity)

# create a dataset where we can extract just the merged data if we want to
diversity_merged <- diversity %>%
  left_join(group_dat)
```

### Check that the merging has worked
```{r}
# check that the merging has worked (row numbers should be equal right now)
nrow(diversity) == nrow(diversity_merged)
```

### Run some tests to see that the merging has worked for all of the sites
```{r}
test_data <- diversity_merged %>%
  filter(nvals_merged > 1) %>%
  distinct(merge_ID, .keep_all = TRUE)

test1 <- diversity_merged %>%
  filter(merge_ID == test_data$merge_ID[1]) %>%
  dplyr::select(SS, SSB,
         Diversity_metric, Diversity_metric_type, Diversity_metric_unit,
         Predominant_habitat, Use_intensity, Years_since_fragmentation_or_conversion,
         Sampling_method, Sampling_effort_unit,
         Study_common_taxon, Rank_of_study_common_taxon,
         Taxon_name_entered,
         Best_guess_binomial,
         Longitude, Latitude,
         Sample_start_earliest, Sample_end_latest, Sample_date_resolution,
         Effort_corrected_measurement,
         Corrected_sampling_effort,
         merged_diversity,
         .rows,
         nvals_merged,
         merge_ID
         )

test1
```
### Run another test (AD2 Marques 2017 needed merging)
```{r}
test2 <- diversity_merged %>%
  filter(merge_ID == test_data$merge_ID[8000]) %>%
  dplyr::select(SS, SSB,
         Diversity_metric, Diversity_metric_type, Diversity_metric_unit,
         Predominant_habitat, Use_intensity, Years_since_fragmentation_or_conversion,
         Sampling_method, Sampling_effort_unit,
         Study_common_taxon, Rank_of_study_common_taxon,
         Taxon_name_entered,
         Best_guess_binomial,
         Longitude, Latitude,
         Sample_start_earliest, Sample_end_latest, Sample_date_resolution,
         Effort_corrected_measurement,
         Corrected_sampling_effort,
         merged_diversity,
         .rows,
         nvals_merged,
         merge_ID
         )

test2
```

### Check studies that did not need merging
```{r}
test_data <- diversity_merged %>%
  filter(nvals_merged == 1) %>%
  distinct(merge_ID, .keep_all = TRUE)

paste(test_data$SS[1])
```

### Run a final test for non-merged sites - test 3
```{r}
test3 <- diversity_merged %>%
  filter(merge_ID == test_data$merge_ID[1]) %>%
  dplyr::select(SS, SSB,
         Diversity_metric, Diversity_metric_type, Diversity_metric_unit,
         Predominant_habitat, Use_intensity, Years_since_fragmentation_or_conversion,
         Sampling_method, Sampling_effort_unit,
         Study_common_taxon, Rank_of_study_common_taxon,
         Taxon_name_entered,
         Best_guess_binomial,
         Longitude, Latitude,
         Sample_start_earliest, Sample_end_latest, Sample_date_resolution,
         Effort_corrected_measurement,
         Corrected_sampling_effort,
         merged_diversity,
         .rows,
         nvals_merged,
         merge_ID
         )

test3
```
### Check how often I am merging values
```{r}
hist(diversity_merged$nvals_merged)
```
### Look at the study where 90 values were merged
This is an extreme case, but all the sites have identical coordinates and land use so the merging was useful
```{r}
diversity_merged %>% 
  filter(nvals_merged == 90)
```

### Subset the merged diversity data into separate dataframes
Create a dataframe with just Dung beetles, just Ants, and a combined dataframe. Since this is before condensing down to just the sites data, I can use these dataframes to get lists of the unique species in the Neotropics, for use when I create a trophic categorization.
```{r}
#rename "Predominant habitat" to "Predominant land use"
diversity_merged <- rename(diversity_merged, Predominant_land_use = Predominant_habitat)

#subset out all other taxa except for insects
insects <- droplevels(subset(diversity_merged, Class=="Insecta"))

#create Neotropical dung beetles dataframe
NeoDungBeetles <- droplevels(subset(insects, Family == "Scarabaeidae" | Family== "Geotrupidae"))
NeoDungBeetles <- droplevels(subset(NeoDungBeetles, Realm == "Neotropic"))

#create Neotropical ants dataframe
NeoAnts <- droplevels(subset(insects, Family == "Formicidae"))
NeoAnts <- droplevels(subset(NeoAnts, Realm == "Neotropic"))

#create dataframe with both in it
NeoAntDungBeetles <- droplevels(subset(insects, Family == "Scarabaeidae" | Family== "Geotrupidae" | Family == "Formicidae"))
NeoAntDungBeetles <- droplevels(subset(NeoAntDungBeetles, Realm == "Neotropic"))
```

### Check if the Scarabaeidae beetles I have are dung beetles for the dung beetles dataset
Create a new dataset of only the Scarabaeidae family, run through the unique genera, and compare this to the NCBI online taxonomy.
```{r}
#Make an only scarab beetle dataset
scarab_check <- droplevels(subset(NeoDungBeetles, Family == "Scarabaeidae"))
 
# Find the unique genus in dataset
scarab_check<- unique(scarab_check$Genus)
 
# Make it a character vector so it is compatible with tax name
scarab_check <- as.character(scarab_check)
 
# Match genus names to a subfamily through ncbi classification and save as dataframe
id <- tax_name(query = c(scarab_check), get = c("subfamily"), db = "ncbi")
 
# Use the get_uid command to search NCBI for the genus names, if you can a taxid (a string of numbers) then it worked, if you get NA then it didn't find anything
uid<-get_uid(c(scarab_check))
 
# Post results into dataframe
uid<-as.data.frame(uid)

# Only need the column that specifies whether there was a match so delete all other columns
uid <- uid[ -c(1,2,4,5,6) ]
 
# Column bind id and uid
ids<-cbind(uid, id)
 
# All of the found ids are dung beetles - all were scarabaeinae.
# Some rows have the family name but no genera information - since there is no genera information I will have to remove the rows that only have the family identified.

 
#Scarab _check is the length of every observation of the family scarabaeidae
scarab_check <- droplevels(subset(NeoDungBeetles, Family == "Scarabaeidae"))
nrow(scarab_check)
#length is  102176
 
#now we just want to take the genus column from only_scarabs and remove any blanks
Genus <- subset(scarab_check, select=c("Genus"))
Genus<-na.omit(Genus)
Genus<- droplevels(subset(Genus, Genus != ""))
nrow(Genus)
#length is 101696
 
102176 - 101696
 
#So there are 480 observations where there is family but no genus classification for scarabaeidae - so I can't determine whether they are dung beetles or not. 
 
test<-NeoDungBeetles[!(NeoDungBeetles$Family=="Scarabaeidae" & NeoDungBeetles$Genus==""),]
 
#test that works
nrow(test)
nrow(NeoDungBeetles)

101725 - 102205 # give 480 - so the method works

## check the table for counts of family
table(test$Family)
```

### Because I will be classifying Ants into trophic groups based on genus accounts, and because dung beetles also need to be identified at the genus level to ensure they are dung beetles, I will remove all rows where genus is unspecified.
```{r}
# Finally, update the dataframes so that Genus does not have any NAs
NeoAnts <- droplevels(subset(NeoAnts, Genus!=""))
NeoDungBeetles <- droplevels(subset(NeoDungBeetles, Genus!=""))
NeoAntDungBeetles <- droplevels(subset(NeoAntDungBeetles, Genus != ""))
```

### Save these as new dataframes for easier loading
```{r}
#create dung beetle file
write.csv(NeoDungBeetles, "NeoDungBeetles.csv")
#create ant file
write.csv(NeoAnts, "NeoAnts.csv")
#create combined file
write.csv(NeoAntDungBeetles, "NeoAntDungBeetles.csv")
```

This is the end of this script