---
title: "2_Combined_DB_Ant_Modelling"
author: "Alex Dhond"
date: "7/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script I will run code to create a combined data-set of ants and dung beetles, and then create linear and generalized linear mixed effects models to see if the abundance and species richness responses differ from one another.

### Clearing workspace prior to running the script
```{r}
rm(list=ls())
```

### Load all required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation

## Modelling ## 
library(lme4) # creating mixed effects models
library(car) # for LRT / anova comparisons
library(merTools) # for extracting model estimates
library(MuMIn) # for assessing model fit
library(optimx) #optimizing models
library(DHARMa) #for model validation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in all required functions
```{r}
## Load in the source function for checking GVIFs - for collinearity
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(1,3))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}

## Load in an overdispersion function from this website to help test if the models are overdispersed.
# https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-for-overdispersioncomputing-overdispersion-factor
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

### Read in the CSV of the combined ant and dung beetles
```{r}
NeoAntDungBeetles <- read.csv("NeoAntDungBeetles.csv")
```

### Create a new column that has the dung beetle families combined as well as ants
Because dung beetles are made up of two families (Geotrupidae and Scarabaeidae) I need to combine the families together.
```{r}
# Make a new column "Family_combined" to put in either Ants or Dung beetles depending on the family
NeoAntDungBeetles <- NeoAntDungBeetles %>%
  mutate(Family_combined = ifelse(Family == "Formicidae", "Ants", "Dung_beetles"))
```

### Create Dung beetle and ant sites - for looking at general land use/LUI/use intensity differences
```{r}
NeoAntDungBeetleSites <- NeoAntDungBeetles %>%

  # pull out only the merged diversity data
  distinct(merge_ID, .keep_all = TRUE) %>%
  
  # re-make SSB and SSBS values 
  mutate(SS = paste(Source_ID, Study_number),
         SSB = paste(SS, Block),
         SSBS = paste(SSB, Site_number)) %>%
  
  # group by SSBS (each unique value corresponds to a unique site)
  group_by(SSBS, Family_combined) %>%
  
  # now add up all the abundance measurements within each site
  mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                 sum(merged_diversity),
                                 # if the diversity metric type isn't Abundance, then leave the TotalAbundance measurement as NA
                                 NA),
         
         # if the metric is already species richness
         SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                  # just use the value as given
                                  merged_diversity,
                                  # for abundance and occurrence measurements, count the number of unique species names that are present at the site 
                                  n_distinct(Taxon_name_entered[merged_diversity > 0])),
         
         # calculate Chao's Species Richness Estimator
         # if the diversity metric is suitable for this calculation
         ChaoRichness = ifelse(Diversity_metric_is_suitable_for_Chao == TRUE,
                               # calculate the Chao estimator
                               sum(merged_diversity > 0) + (((sum(merged_diversity == 1) * (sum(merged_diversity == 1)-1)) / (2*(sum(merged_diversity == 2)+1)))),
                               # otherwise give Chao NA
                               NA)
         ) %>%
  
  # ungroup
  ungroup() %>%
  
    # now group by Study ID
  group_by(SS, Family_combined) %>%
  
  # pull out some useful study-level numbers
  # maximum abundance for each study
  mutate(MaxAbundance = max(TotalAbundance),
         # minimum (non-zero) abundance for each study
         # for species rarefaction
         MinNonZeroAbundance = min(TotalAbundance[TotalAbundance > 0]),
         # number of species in the study
         SpeciesInStudy = n_distinct(Taxon_name_entered[merged_diversity > 0]),
         # assess whether the study is suitable for rarefaction
         SuitableForRarefaction = ifelse(
           # if all diversity measurements are integers
           # i.e. if you round down, it should be equal to the original number
           # and the diversity metric is suitable for Chao
           all(floor(merged_diversity) == merged_diversity) &
             Diversity_metric_is_suitable_for_Chao == TRUE,
           # then class the study as suitable for rarefaction
           TRUE,
           # otherwise it can't be used
           FALSE
           )
         ) %>%
  
  # ungroup
  ungroup()  %>%
  
  # now rescale total abundance, so that within each study, abundance varies from 0 to 1.
  mutate(RescaledAbundance = TotalAbundance/MaxAbundance,
         # for statistical modelling,  calculate the square root of species abundance, although might want to use log(x+1) transformation instead
         sqrtRescaledAbundance = sqrt(RescaledAbundance),
         logRescaledAbundance = log(RescaledAbundance + 1)
         )

# pull out unique sites
NeoAntDungBeetleSites <- NeoAntDungBeetleSites %>%
  distinct(SSBS, Family_combined, .keep_all = TRUE)
```

### Look at the site data, and collapse variables with little/no data
```{r}
NeoAntDungBeetleSites <- NeoAntDungBeetleSites %>%
  
  mutate(

    #collapse primary vegetation together
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                  "Primary forest" = "Primary vegetation", 
                                  "Primary non-forest" = "Primary vegetation"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Use_intensity = na_if(Use_intensity, "Cannot decide"),

  )

#remove urban sites
NeoAntDungBeetleSites <- droplevels(subset(NeoAntDungBeetleSites, Predominant_land_use != "Urban"))

# take another look at the LandUse/Use intensity split
table(NeoAntDungBeetleSites$Predominant_land_use, NeoAntDungBeetleSites$Use_intensity, NeoAntDungBeetleSites$Family_combined) 
```

Due to imbalances in the data, I can only look at predominant land use, and not the more coarse LUI, when comparing ants and dung beetles. I want to have at least 30 observations in each category (for both groups of species) so unfortunately I cannot do a finer LUI classification.

### Rename the land use classifications so it is easier to read
First revalue/relabel the land use categories so that graphs are easier to read. Then recode the levels so that when modelling they appear with primary vegetation as the reference level (and on graphs they go in an intuitive order)
```{r}
## Revalue the factors so they go in logical order
NeoAntDungBeetleSites <- NeoAntDungBeetleSites %>%
  mutate(Predominant_land_use = plyr::revalue(Predominant_land_use, c("Primary vegetation" = "Primary",
                                    "Mature secondary vegetation" = "MSV",
                                    "Intermediate secondary vegetation" = "ISV",
                                    "Young secondary vegetation" = "YSV",
                                    "Plantation forest" = "Plantation",
                                    "Pasture" = "Pasture",
                                    "Cropland" = "Cropland")))

## Check to see the names are new
table(NeoAntDungBeetleSites$Predominant_land_use)

## Now re-code the levels
levels(NeoAntDungBeetleSites$Predominant_land_use)

#recode the levels
NeoAntDungBeetleSites$Predominant_land_use <- factor(NeoAntDungBeetleSites$Predominant_land_use, levels = c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland"))

#check it worked
levels(NeoAntDungBeetleSites$Predominant_land_use)
```

### check the number of sources sites and studies
```{r}
length(unique(NeoAntDungBeetleSites$Source_ID)) #number of sources
length(unique(NeoAntDungBeetleSites$SS)) #number of studies
length(unique(NeoAntDungBeetleSites$SSBS)) #number of sites
```

The combined data should be in order now for modelling. Now create the model data (need to drop out any NAs that would interfere with the models being run).

For the model data, I am going to model abundance (log and sqrt transformed) as well as Species richness.

## Create the combined dataset
```{r}
## Create the combined abundance dataframe, dropping NAs from the relevant columns
model_data_comb_ab <- drop_na(NeoAntDungBeetleSites,
                              RescaledAbundance, Predominant_land_use, Family_combined)

# Add in the transformed log and sqrt abundances
model_data_comb_ab <- mutate(model_data_comb_ab, 
                        logAbundance = log(RescaledAbundance + 1),
                        sqrtAbundance = sqrt(RescaledAbundance))

## Create the combined Species richness dataframe, dropping out NAs
model_data_comb_sr <- drop_na(NeoAntDungBeetleSites,
                              SpeciesRichness, Predominant_land_use, Family_combined)

## In case of overdispersion, add in a row-level random effect
model_data_comb_sr$Row_random_effect <- 1:nrow(model_data_comb_sr)
```

Now that the model dataframes have been created, I can run models on transformed abundances for the combined dataset. I will start with species richness

## SPECIES RICHNESS - check interaction between land use and dung beetle and ants

### First check for collinearity between the predictor variables
I don't expect there to be any collinearity between predominant land use and Family_combined, but I will check anyway
```{r}
## Test for collinearity
corvif(NeoAntDungBeetleSites[ , c("Predominant_land_use", "Family_combined")])
```

### Create models and proceed with backwards stepwise selection
```{r}
## Create the maximal model
m1 <- glmer(SpeciesRichness ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Remove random effects to test 
m2 <- glmer(SpeciesRichness ~ Predominant_land_use*Family_combined + (1|SS), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Remove study
m3 <- glmer(SpeciesRichness ~ Predominant_land_use*Family_combined + (1|SSB), 
            data = model_data_comb_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Check the AIC of the models to see which random effects structure has the lowest AIC
AIC(m1, m2, m3)

## Model m1 is the best random effect structure

## Proceed with backwards stepwise selection of fixed effects

## Remove the interaction term but keep the family_combined variable
m4 <- glmer(SpeciesRichness ~ Predominant_land_use + Family_combined + (1|SS) + (1|SSB), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Remove the family combined variable
m5 <- glmer(SpeciesRichness ~ Predominant_land_use + (1|SS) + (1|SSB), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Remove predominant land use
m6 <- glmer(SpeciesRichness ~ Family_combined + (1|SS) + (1|SSB), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Create a null model with no fixed effects
m7 <- glmer(SpeciesRichness ~ (1|SS) + (1|SSB), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Compare the models
anova(m1, m4, m5, m6, m7)
model.sel(m1, m4, m5, m6, m7)
AIC(m1, m4, m5, m6, m7)

## Model m1 is the best model with the lowest AIC

## Check anova
Anova(m1)
```

The maximal model, m1, is the most significant model with the lowest AIC. I will now test it for overdispersion.

### Test the model for overdispersion
```{r}
## Use the overdisp function
overdisp_fun(m1)
```

Because the p-value is significant, the model is overdispersed. I will now add in another random effect to account for overdispersion

### Add in another random effect to account for over dispersion
```{r}
## Add each row (essentially site) as a random effect
m8 <- glmer(SpeciesRichness ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_data_comb_sr,
            family = "poisson",
            control = glmerControl(optimizer = "bobyqa"))

## Check for overdispersion
overdisp_fun(m8)
```

The p-value for m8 overdispersion test is not significant, so model m8 should be acceptable to model with.

### Check DHARMa model diagnostics for species richness model
```{r}
# look at the model criticism plots
model_plot(m8)
## Simulate the residuals of the model using DHARMa
m8_residuals <- simulateResiduals(fittedModel = m8, plot = F)

## Plot them
plot(m8_residuals)

## Test the outliers and test for dispersion
testOutliers(m8_residuals)
testDispersion(m8_residuals)
```

Evidently my model has some outliers. I have tested for overdispersion, and the model is not over-dispersed. 

### Before I move on, get the summary statistics and anova values, as well as any other relevant information for the methods section
```{r}
## First, make a model to compare the interaction term between land use and family combined
m4_test <- glmer(SpeciesRichness ~ Predominant_land_use + Family_combined + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_data_comb_sr,
            family = "poisson",
            )

## Use anova to check if the interaction is significant
anova(m4_test, m8)

## Comparing the model m4a_test to m8, the interaction between land use and family was significant. Chi squared = 34.12, df = 6, p  = 6.377e-06 *** (copy this to methods)

## get the summary statistics
summary(m8)
```

# ABUNDANCE
## Make the sqrt abundance models 

### Create sqrt abundance models and proceed with backwards stepwise selection
```{r}
## Start with removing the random effects 

## Create the maximal model
m1a <- lmer(sqrtAbundance ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Remove block
m2a <- lmer(sqrtAbundance ~ Predominant_land_use*Family_combined + (1|SS), data = model_data_comb_ab)

## Remove study, keep block
m3a <- lmer(sqrtAbundance ~ Predominant_land_use*Family_combined + (1|SSB), data = model_data_comb_ab)

## Compare the models for the best random effects structure
AIC(m1a, m2a, m3a)

## Model m1a has the lowest AIC, so that random effects structure is the best

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
m4a<- lmer(sqrtAbundance ~ Predominant_land_use + Family_combined + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Sqrt abundance model with just land use as predictor
m5a <- lmer(sqrtAbundance ~ Predominant_land_use + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Model with just family_combined as the predictor
m6a <- lmer(sqrtAbundance ~ Family_combined + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Make a null, intercept only model
m7a <- lmer(sqrtAbundance ~ (1|SS) + (1|SSB), data = model_data_comb_ab)


## Compare the models via anova and model.sel
anova(m1a, m4a, m5a, m6a, m7a)
model.sel(m1a, m4a, m5a, m6a, m7a)
AIC(m1a, m4a, m5a, m6a, m7a)

anova(m1a, m4a)

## From these tests, it seems that model m1a is the best
```

For the sqrt abundance models, model m1a is the best model

### Check model diagnostics for model m1a
```{r}
## Check the residuals from the model criticism function
model_plot(m1a)

## Simulate the residuals 
m1a_residuals <- simulateResiduals(fittedModel = m6a, plot = F)

## Plot them
plot(m1a_residuals)

testOutliers(m1a_residuals)
testDispersion(m1a_residuals)

## Check for overdispersion with the loaded function
overdisp_fun(m1a)
```

The sqrt abundance model, m1a, shows relatively normal distribution of residuals. It seems to be pretty good

Move onto log abundance.

## Create the log abundance models
```{r}
## Start with removing the random effects 

## Create the maximal model
m1a1 <- lmer(logAbundance ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Remove block
m2a1 <- lmer(logAbundance ~ Predominant_land_use*Family_combined + (1|SS), data = model_data_comb_ab)

## Remove study, keep block
m3a1 <- lmer(logAbundance ~ Predominant_land_use*Family_combined + (1|SSB), data = model_data_comb_ab)

## Compare the models for the best random effects structure
AIC(m1a1, m2a1, m3a1)

## Model m1a has the lowest AIC, so that random effects structure is the best

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
m4a1 <- lmer(logAbundance ~ Predominant_land_use + Family_combined + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Sqrt abundance model with just land use as predictor
m5a1 <- lmer(logAbundance ~ Predominant_land_use + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Model with just family_combined as the predictor
m6a1 <- lmer(logAbundance ~ Family_combined + (1|SS) + (1|SSB), data = model_data_comb_ab)

## Make a null, intercept only model
m7a1 <- lmer(logAbundance ~ (1|SS) + (1|SSB), data = model_data_comb_ab)


## Compare the models via anova and model.sel
anova(m1a1, m4a1, m5a1, m6a1, m7a1)
model.sel(m1a1, m4a1, m5a1, m6a1, m7a1)
AIC(m1a1, m4a1, m5a1, m6a1, m7a1)

## Compare the best two models
anova(m1a1, m4a1)

## From these tests, it seems that model m1a1 is the best
```

### Check the summary statistics and diagnostic plots for this model
```{r}
## Check the model diagnostics
model_plot(m1a1)

## Simulate the residuals 
m1a1_residuals <- simulateResiduals(fittedModel = m1a1, plot = F)

## Plot them
plot(m1a1_residuals)

## Once again the plots are extremely skewed
testDispersion(m1a1_residuals)

## Test for overdispersion
overdisp_fun(m1a1)
```

The log abundance model, m1a1, shows a lot more skew than the sqrt abundance model. As such, I will probably use the sqrt abundance model as my go to model.




# PLOTTING AND VISUALIZATION

# Visualize the data

### Graph sqrt and log abundance
```{r}
## Sqrt abundance graph
ggplot(model_data_comb_ab, aes(x = Predominant_land_use, y = sqrtAbundance, colour = Family_combined))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45))

## Log abundance graph
ggplot(model_data_comb_ab, aes(x = Predominant_land_use, y = logAbundance, colour = Family_combined))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45))

## Species richness graph
ggplot(model_data_comb_sr, aes(x = Predominant_land_use, y = SpeciesRichness, colour = Family_combined))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45))+
  theme(legend.position = "none")
```

Now I will generate the effects of the combined SR model to make a figure and plot how ants and dung beetles are responding differently. Here, I will need to create two dataframes, one with only the dung beetle effects, and one with only the ant effects. I will transform the effects into percentage differences, then combine the dataframes into one for plotting

## Create the dung beetle effects dataframe

### First create a new model with dung beetles as the reference category
```{r}
## Make the "Family_combined" column a factor
model_data_comb_sr$Family_combined <- as.factor(model_data_comb_sr$Family_combined)

## Create model dataframe with dung beetle as reference category
model_db_sr <- within(model_data_comb_sr, Family_combined <- relevel(Family_combined, ref = "Dung_beetles"))

## Create model 8 with dung beetles
m8_db <- glmer(SpeciesRichness ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_db_sr,
            family = "poisson",
            control = glmerControl("bobyqa"))
```

### Check effects for methods
```{r}
summary(m8_db)
summary(m)
```

### Simulate the db model effects
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_eff <- FEsim(m8_db, 1000)

# make the term column a factor so it can be recoded
db_eff$term <- as.factor(db_eff$term)

# pull out dung beetle effects
db_eff <- db_eff %>%
  slice(1:7)

# recode the factors to make them shorter (so I can visualise easier)
db_eff$term <- recode_factor(db_eff$term, "(Intercept)" = "Primary",
                              "Predominant_land_useMSV" = "MSV",
                              "Predominant_land_useISV" = "ISV",
                              "Predominant_land_useYSV" = "YSV",
                              "Predominant_land_usePlantation" = "Plantation",
                              "Predominant_land_usePasture" = "Pasture",
                              "Predominant_land_useCropland" = "Cropland")

# Add in upper and lower confidence intervals 
db_eff <- db_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the estimates and generate percentage differences from the baseline (first term, Primary Vegetation (PV))
db_eff <- db_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
db_eff[1,7] <- 0 #shift median
db_eff[1,8] <- 0 #shift upper
db_eff[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(db_eff)[1] <- "Predominant_land_use"
```

### Create the ant effects dataframe
```{r}
# gather the effects and confidence intervals using simulation. Simulate 10000 times
ant_eff <- FEsim(m8, 1000)

# make the term column a factor so it can be recoded
ant_eff$term <- as.factor(ant_eff$term)

# Pull out only ant terms
ant_eff <- ant_eff %>%
  slice(1:7)

# recode the factors to make them shorter (so I can visualise easier)
ant_eff$term <- recode_factor(ant_eff$term, "(Intercept)" = "Primary",
                              "Predominant_land_useMSV" = "MSV",
                              "Predominant_land_useISV" = "ISV",
                              "Predominant_land_useYSV" = "YSV",
                              "Predominant_land_usePlantation" = "Plantation",
                              "Predominant_land_usePasture" = "Pasture",
                              "Predominant_land_useCropland" = "Cropland")

# Add in upper and lower confidence intervals 
ant_eff <- ant_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the estimates and generate percentage differences from the baseline (first term, Primary Vegetation (PV)
ant_eff <- ant_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
ant_eff[1,7] <- 0 #shift median
ant_eff[1,8] <- 0 #shift upper
ant_eff[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(ant_eff)[1] <- "Predominant_land_use"
```

### Combine the dataframes together
Add in a column of "Family_combined" to make sure there is a grouping variable
```{r}
comb_sr_eff <- rbind(db_eff, ant_eff)
comb_sr_eff$Family_combined <- c("Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles","Ants", "Ants", "Ants", "Ants", "Ants", "Ants", "Ants")
```

### Finally, plot the combined effect estimates

### Create necessary vectors and objects for plotting
```{r}
## Make grouping colors
color <- c("#00BFC4", "#F8766D")

## set the x axis limits
comb_plot_limits <- c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland")
```

### Plot the percentage differences - SPECIES RICHNESS
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
species_richness <- comb_sr_eff %>%
  ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, colour = Family_combined)+
  geom_hline(yintercept = 0, size = 0.5, color = ("black"), linetype = 1)+
  geom_point(size = 3, position = position_dodge(width = 0.5))+
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), size = 1)+
  theme_classic()+
  scale_x_discrete(limits=comb_plot_limits)+
  geom_vline(xintercept = 1.5, linetype = 2)+
  geom_vline(xintercept = 2.5, linetype = 2)+
  geom_vline(xintercept = 3.5, linetype = 2)+
  geom_vline(xintercept = 4.5, linetype = 2)+
  geom_vline(xintercept = 5.5, linetype = 2)+
  geom_vline(xintercept = 6.5, linetype = 2)+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none",
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Species richness difference (%)")+
  scale_colour_manual(legend_title, values = color)
species_richness
```

From the figure, it is clear that across most land-use types, ant and dung beetle species richness responds similarly, except in mature secondary vegetation and cropland. N = 2086 observations

# ABUNDANCE EFFECTS VISUALIZATION

## Create the dung beetle effects dataframe

### First create a new model with dung beetles as the reference category
```{r}
## Make the "Family_combined" column a factor
model_data_comb_ab$Family_combined <- as.factor(model_data_comb_ab$Family_combined)

## Create model dataframe with dung beetle as reference category
model_db_ab <- within(model_data_comb_ab, Family_combined <- relevel(Family_combined, ref = "Dung_beetles"))

## Create log abundance model
m1a1_db <- lmer(logAbundance ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB), data = model_db_ab)

## sqrt abundance model
m1a_db <- lmer(sqrtAbundance ~ Predominant_land_use*Family_combined + (1|SS) + (1|SSB), data = model_db_ab)
```

### Make db effects for LOG ABUNDANCE models
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_eff_ab <- FEsim(m1a1_db, 1000)

# make the term column a factor so it can be recoded
db_eff_ab$term <- as.factor(db_eff_ab$term)

# pull out dung beetle effects
db_eff_ab <- db_eff_ab %>%
  slice(1:7)

# recode the factors to make them shorter (so I can visualise easier)
db_eff_ab$term <- recode_factor(db_eff_ab$term, "(Intercept)" = "Primary",
                              "Predominant_land_useMSV" = "MSV",
                              "Predominant_land_useISV" = "ISV",
                              "Predominant_land_useYSV" = "YSV",
                              "Predominant_land_usePlantation" = "Plantation",
                              "Predominant_land_usePasture" = "Pasture",
                              "Predominant_land_useCropland" = "Cropland")

# Add in upper and lower confidence intervals 
db_eff_ab <- db_eff_ab %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the log abundance estimates so that I get a measure of abundance
db_eff_ab <- db_eff_ab %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

# Shift the baseline down to 0
db_eff_ab[1,7] <- 0 #shift median
db_eff_ab[1,8] <- 0 #shift upper
db_eff_ab[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(db_eff_ab)[1] <- "Predominant_land_use"
```

### Create the ant effects dataframe
```{r}
# gather the effects and confidence intervals using simulation. Simulate 10000 times
ant_eff_ab <- FEsim(m1a, 1000)

# make the term column a factor so it can be recoded
ant_eff_ab$term <- as.factor(ant_eff_ab$term)

# Pull out only ant terms
ant_eff_ab <- ant_eff_ab %>%
  slice(1:7)

# recode the factors to make them shorter (so I can visualise easier)
ant_eff_ab$term <- recode_factor(ant_eff_ab$term, "(Intercept)" = "Primary",
                              "Predominant_land_useMSV" = "MSV",
                              "Predominant_land_useISV" = "ISV",
                              "Predominant_land_useYSV" = "YSV",
                              "Predominant_land_usePlantation" = "Plantation",
                              "Predominant_land_usePasture" = "Pasture",
                              "Predominant_land_useCropland" = "Cropland")

# Add in upper and lower confidence intervals 
ant_eff_ab <- ant_eff_ab %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the log abundance estimates so that I get a measure of abundance
ant_eff_ab <- ant_eff_ab %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

# Shift the baseline down to 0
ant_eff_ab[1,7] <- 0 #shift median
ant_eff_ab[1,8] <- 0 #shift upper
ant_eff_ab[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(ant_eff_ab)[1] <- "Predominant_land_use"
```

### Combine the dataframes together
Add in a column of "Family_combined" to make sure there is a grouping variable
```{r}
comb_log_ab <- rbind(db_eff_ab, ant_eff_ab)
comb_log_ab$Family_combined <- c("Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles","Ants", "Ants", "Ants", "Ants", "Ants", "Ants", "Ants")
```

### Finally, plot the combined effect estimates
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
comb_log_ab %>%
  ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, colour = Family_combined)+
  geom_point(size = 2, position = position_dodge(width = 0.5))+
  geom_errorbar(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), size = 0.75, width = 0.2)+
  geom_hline(yintercept = 0, size = I(1), color = I("black"), linetype = 1 )+
  theme_light()+
  theme(axis.text.x = element_text(colour = c("#000000", "#E69F00", "#56B4E9", "#009E73",  "#F0E442", "#0072B2","#D55E00"),face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"))+
  xlab("Land use intensity class")+
  ylab("Total abundance difference (%)")+
  scale_colour_manual(legend_title, values = color)
```
This is the result of the log abundance model, back transformed. I dont want to use this graph

Figure x: Total abundance difference from primary vegetation for ants and dung beetles. 


```{r}
summary(m1a_db)
```

### Do the same for the sqrt abundance model
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_eff_sqrt_ab <- FEsim(m1a_db, 1000)

# make the term column a factor so it can be recoded
db_eff_sqrt_ab$term <- as.factor(db_eff_sqrt_ab$term)

# pull out dung beetle effects
db_eff_sqrt_ab <- db_eff_sqrt_ab %>%
  slice(1:7)

# recode the factors to make them shorter (so I can visualise easier)
db_eff_sqrt_ab$term <- recode_factor(db_eff_sqrt_ab$term, "(Intercept)" = "Primary",
                              "Predominant_land_useMSV" = "MSV",
                              "Predominant_land_useISV" = "ISV",
                              "Predominant_land_useYSV" = "YSV",
                              "Predominant_land_usePlantation" = "Plantation",
                              "Predominant_land_usePasture" = "Pasture",
                              "Predominant_land_useCropland" = "Cropland")

# Add in upper and lower confidence intervals 
db_eff_sqrt_ab <- db_eff_sqrt_ab %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the sqrt abundance estimates so that I get a measure of abundance
db_eff_sqrt_ab <- db_eff_sqrt_ab %>%
  mutate(
    Percent_diff = ((((median[1] + median)^2) / (median[1])^2)*100)-100) %>%
  mutate(
    Percent_upper = (((median[1] + Upper_ci)^2) / ((median[1])^2) * 100) - 100) %>%
  mutate(
    Percent_lower = ((((median[1] + Lower_ci)^2) / ((median[1])^2) * 100) - 100)
  )

# Shift the baseline down to 0
db_eff_sqrt_ab[1,7] <- 0 #shift median
db_eff_sqrt_ab[1,8] <- 0 #shift upper
db_eff_sqrt_ab[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(db_eff_sqrt_ab)[1] <- "Predominant_land_use"
```

### Create the ant effects dataframe
```{r}
# gather the effects and confidence intervals using simulation. Simulate 10000 times
ant_eff_sqrt_ab <- FEsim(m1a, 1000)

# make the term column a factor so it can be recoded
ant_eff_sqrt_ab$term <- as.factor(ant_eff_sqrt_ab$term)

# Pull out only ant terms
ant_eff_sqrt_ab <- ant_eff_sqrt_ab %>%
  slice(1:7)

# recode the factors to make them shorter (so I can visualise easier)
ant_eff_sqrt_ab$term <- recode_factor(ant_eff_sqrt_ab$term, "(Intercept)" = "Primary",
                              "Predominant_land_useMSV" = "MSV",
                              "Predominant_land_useISV" = "ISV",
                              "Predominant_land_useYSV" = "YSV",
                              "Predominant_land_usePlantation" = "Plantation",
                              "Predominant_land_usePasture" = "Pasture",
                              "Predominant_land_useCropland" = "Cropland")

# Add in upper and lower confidence intervals 
ant_eff_sqrt_ab <- ant_eff_sqrt_ab %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the sqrt abundance estimates so that I get a measure of abundance
ant_eff_sqrt_ab <- ant_eff_sqrt_ab %>%
  mutate(
    Percent_diff = ((((median[1] + median)^2) / (median[1])^2)*100)-100) %>%
  mutate(
    Percent_upper = (((median[1] + Upper_ci)^2) / ((median[1])^2) * 100) - 100) %>%
  mutate(
    Percent_lower = ((((median[1] + Lower_ci)^2) / ((median[1])^2) * 100) - 100)
  )

# Shift the baseline down to 0
ant_eff_sqrt_ab[1,7] <- 0 #shift median
ant_eff_sqrt_ab[1,8] <- 0 #shift upper
ant_eff_sqrt_ab[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(ant_eff_sqrt_ab)[1] <- "Predominant_land_use"
```

### Combine the dataframes together
Add in a column of "Family_combined" to make sure there is a grouping variable
```{r}
comb_sqrt_ab <- rbind(db_eff_sqrt_ab, ant_eff_sqrt_ab)
comb_sqrt_ab$Family_combined <- c("Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles", "Dung beetles","Ants", "Ants", "Ants", "Ants", "Ants", "Ants", "Ants")
```

### Finally, plot the combined effect estimates
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
abundance <- comb_sqrt_ab %>%
  ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, colour = Family_combined)+
  geom_hline(yintercept = 0, size = 0.5, color = ("black"), linetype = 1)+
  geom_point(size = 3, position = position_dodge(width = 0.5))+
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), size = 1)+
  theme_classic()+
  scale_x_discrete(limits=comb_plot_limits)+
  geom_vline(xintercept = 1.5, linetype = 2)+
  geom_vline(xintercept = 2.5, linetype = 2)+
  geom_vline(xintercept = 3.5, linetype = 2)+
  geom_vline(xintercept = 4.5, linetype = 2)+
  geom_vline(xintercept = 5.5, linetype = 2)+
  geom_vline(xintercept = 6.5, linetype = 2)+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        legend.position = "none",
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Total abundance difference (%)")+
  scale_colour_manual(legend_title, values = color)
abundance
```
Figure X: Linear mixed effects model showing total abundance difference from primary vegetation. Error bars show the 95% confidence interval around each estimate. Error bars that do not cross zero are statistically significantly different from the baseline (primary vegetation). Line colors represent the following taxonomic groups: teal (Ants), red (Dung beetles). X-axis colors represent the following land-use types: black (primary vegetation, Primary), orange (mature secondary vegetation, MSV), light blue (intermediate secondary vegetation, ISV), green (young secondary vegetation, YSV), yellow (plantation forest, Plantation), dark blue (Pasture), dark orange (Cropland). N = 1816 observations

### Make the legend
```{r}
### Make a new plot, extract just the legend from it,so I can combine with my other nicer plots
comb_legend_1 <- ggplot(model_data_comb_sr, aes(x = Predominant_land_use, fill = Family_combined)) + 
  scale_fill_manual(labels = c("Ants", "Dung beetles"), values=color)+
  guides(fill=guide_legend(title = "Taxonomic group", title.position = "top", nrow=1, byrow=TRUE))+
  theme(legend.position = "bottom",
        legend.text = element_text( face = "bold", size = 8),
        legend.title = element_text(face = "bold"))+
  geom_bar()
comb_legend_1

# Using the cowplot package
comb_legend_2 <- cowplot::get_legend(comb_legend_1)
comb_plot_legend <- as_ggplot(comb_legend_2)
comb_plot_legend
```

### Export the graphs
```{r}
combined_diversity_plots_1 <- ggarrange(abundance, species_richness,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2)
combined_diversity_plots_2 <- ggarrange(abundance, comb_plot_legend,
                                        labels = c("", ""),
                                        ncol = 1, nrow = 2)
combined_diversity_plots_2
ggexport(combined_diversity_plots_1, filename = "combined_diversity_plots_1.pdf")
ggexport(combined_diversity_plots_2, filename = "combined_diversity_plots_2.pdf")
```
