---
title: "3_DB_site_diversity"
author: "Alex Dhond"
date: "7/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for testing Dung beetle site biodiversity metrics on their own. In the combined dataset, I was not able to get a fine scale look at the Land use intensity classes due to imbalances between ants and dung beetles. In this script, I want to model dung beetle abundance and species richness as a response to land use intensity (which I will create as a new variable).

Because the Simpson's diversity requires me to make a matrix, I have done that in a separate script (script 3a), and the modelling for Simpson's index in script 3b. Since I will be plotting Simpson's together with the abundance and species richness, I have also made a separate script for the plotting (see script 3c).

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation

## Modelling ## 
library(lme4) # creating mixed effects models
library(car) # for LRT / anova comparisons
library(merTools) # for extracting model estimates
library(MuMIn) # for assessing model fit
library(optimx) #optimizing models
library(DHARMa) #for model validation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in any required functions
```{r}
## Load in the source function for checking GVIFs - for collinearity
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(1,3))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}

## Load in an overdispersion function from this website to help test if the models are overdispersed.
# https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-for-overdispersioncomputing-overdispersion-factor
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

### Load in a colour palette (for the figures)
```{r}
## Load in a color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### First read in the Dung beetle data
```{r}
NeoDungBeetles <- read.csv("NeoDungBeetles.csv")
```

# Calculate broad site level diversity metrics - abundance and species richness
The Dung beetle data in the Neotropics is mainly abundance data, so I can calculate site level abundance and species richness estimates

### Create Neo dung beetle sites
```{r}
NeoDungBeetleSites <- NeoDungBeetles %>%

  # pull out only the merged diversity data
  distinct(merge_ID, .keep_all = TRUE) %>%
  
  # re-make SSB and SSBS values since we've now dropped a bunch of values
  mutate(SS = paste(Source_ID, Study_number),
         SSB = paste(SS, Block),
         SSBS = paste(SSB, Site_number)) %>%
  
  # group by SSBS (each unique value corresponds to a unique site)
  group_by(SSBS) %>%
  
  # now add up all the abundance measurements within each site
  mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                 sum(merged_diversity),
                                 # if the diversity metric type isn't Abundance, then leave the TotalAbundance measurement as NA
                                 NA),
         
         # if the metric is already species richness
         SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                  # just use the value as given
                                  merged_diversity,
                                  # for abundance and occurrence measurements, count the number of unique species names that are present at the site 
                                  n_distinct(Taxon_name_entered[merged_diversity > 0])),
         
         # calculate Chao's Species Richness Estimator
         # if the diversity metric is suitable for this calculation
         ChaoRichness = ifelse(Diversity_metric_is_suitable_for_Chao == TRUE,
                               # calculate the Chao estimator
                               sum(merged_diversity > 0) + (((sum(merged_diversity == 1) * (sum(merged_diversity == 1)-1)) / (2*(sum(merged_diversity == 2)+1)))),
                               # otherwise give Chao NA
                               NA)
         ) %>%
  
  # ungroup
  ungroup() %>%
  
    # now group by Study ID
  group_by(SS) %>%
  
  # pull out some useful study-level numbers
  # maximum abundance for each study
  mutate(MaxAbundance = max(TotalAbundance),
         # minimum (non-zero) abundance for each study
         # we'll use this when we do species rarefaction
         MinNonZeroAbundance = min(TotalAbundance[TotalAbundance > 0]),
         # number of species in the study
         SpeciesInStudy = n_distinct(Taxon_name_entered[merged_diversity > 0]),
         # assess whether the study is suitable for rarefaction
         SuitableForRarefaction = ifelse(
           # if all diversity measurements are integers
           # i.e. if you round down, it should be equal to the original number
           # and the diversity metric is suitable for Chao
           all(floor(merged_diversity) == merged_diversity) &
             Diversity_metric_is_suitable_for_Chao == TRUE,
           # then class the study as suitable for rarefaction
           TRUE,
           # otherwise it can't be used
           FALSE
           )
         ) %>%
  
  # ungroup
  ungroup()  %>%
  
  # now rescale total abundance, so that within each study, abundance varies from 0 to 1.
  mutate(RescaledAbundance = TotalAbundance/MaxAbundance,
         # for statistical modelling, we'll also calculate the square root of species abundance, although we might want to use log(x+1) transformation instead
         sqrtRescaledAbundance = sqrt(RescaledAbundance),
         logRescaledAbundance = log(RescaledAbundance + 1)
         )

# pull out unique sites
NeoDungBeetleSites <- NeoDungBeetleSites %>%
  distinct(SSBS, .keep_all = TRUE)
```

### check the number of sources, studies, and sites
```{r}
#dung beetles
length(unique(NeoDungBeetleSites$Source_ID)) #number of sources
length(unique(NeoDungBeetleSites$SS)) #number of studies
length(unique(NeoDungBeetleSites$SSBS)) #number of sites
```

### Look at the site data, and collapse variables with little/no data
Here I am removing collapsing primary vegetation, removing the indeterminate as well as cannot decide categories, and removing urban as a category
```{r}
NeoDungBeetleSites <- NeoDungBeetleSites %>%
  
  mutate(

    #collapse primary vegetation together
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                  "Primary forest" = "Primary vegetation", 
                                  "Primary non-forest" = "Primary vegetation"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Use_intensity = na_if(Use_intensity, "Cannot decide"),

  )

#remove urban sites
NeoDungBeetleSites <- droplevels(subset(NeoDungBeetleSites, Predominant_land_use != "Urban"))

# take another look at the LandUse/Use intensity split
table(NeoDungBeetleSites$Predominant_land_use, NeoDungBeetleSites$Use_intensity) 
```

### Collapse all data in which there are 30 or less observations per land use/intensity combination.
30 observations will be the cut off point. Here I am collapsing land use intensities together to make sure there is an even enough data spread.
```{r}
NeoDungBeetleSites <- NeoDungBeetleSites %>%
  mutate(LUI = interaction(Predominant_land_use, Use_intensity, sep = "_"),
         
         #collapse primary vegetation Light Intense together
         LUI = recode_factor(LUI, "Primary vegetation_Light use" = "Primary vegetation_LightIntense", "Primary vegetation_Intense use" = "Primary vegetation_LightIntense"),
         
         #collapse Cropland Intense and Light together, save as "LightIntense"
         LUI = recode_factor(LUI, "Cropland_Intense use" = "Cropland_LightIntense",
                             "Cropland_Light use" = "Cropland_LightIntense"),
         
         #collapse Mature Secondary Vegetation Intense and Light into "LightIntense"
         LUI = recode_factor(LUI, 
                             "Mature secondary vegetation_Minimal use" = "Mature secondary vegetation",
                             "Mature secondary vegetation_Intense use" = "Mature secondary vegetation", 
                             "Mature secondary vegetation_Light use" = "Mature secondary vegetation"),
         
         #collapse plantation forest intense and light together
         LUI = recode_factor(LUI, "Plantation forest_Intense use" = "Plantation forest",
                             "Plantation forest_Light use" = "Plantation forest",
                             "Plantation forest_Minimal use" = "Plantation forest"),
         
         #collapse Intermediate secondary vegetation light and intense together
         LUI = recode_factor(LUI, "Intermediate secondary vegetation_Intense use" = "Intermediate secondary vegetation", "Intermediate secondary vegetation_Light use" = "Intermediate secondary vegetation",
                             "Intermediate secondary vegetation_Minimal use" = "Intermediate secondary vegetation"),
         
         #collapse Young secondary vegetation light and intense together
         LUI = recode_factor(LUI, "Young secondary vegetation_Light use" = "Young secondary vegetation", "Young secondary vegetation_Intense use" = "Young secondary vegetation",
                             "Young secondary vegetation_Minimal use" = "Young secondary vegetation"),

         #dont collapse pasture
  )

table(NeoDungBeetleSites$LUI)
table(NeoDungBeetleSites$Predominant_land_use, NeoDungBeetleSites$Use_intensity)
```

### Recode these factor levels for easier visualization on graphs
```{r}
#LightIntense use category will be classified as "Higher (H)" use for brevity
NeoDungBeetleSites <- NeoDungBeetleSites %>%
  mutate(LUI = plyr::revalue(LUI, c("Primary vegetation_Minimal use" = "PVMU",
                                    "Primary vegetation_LightIntense" = "PVHU",
                                    "Mature secondary vegetation" = "MSV",
                                    "Intermediate secondary vegetation" = "ISV",
                                    "Young secondary vegetation" = "YSV",
                                    "Plantation forest" = "PF",
                                    "Pasture_Minimal use" = "PMU",
                                    "Pasture_Light use" = "PLU",
                                    "Pasture_Intense use" = "PIU",
                                    "Cropland_Minimal use" = "CMU",
                                    "Cropland_LightIntense" = "CHU")))
```

### Reorder factor levels so that PVMU is the baseline for modelling (and graphing)
```{r}
#first get a list of the levels
levels(NeoDungBeetleSites$LUI)

#recode the levels
NeoDungBeetleSites$LUI <- factor(NeoDungBeetleSites$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#check it worked
levels(NeoDungBeetleSites$LUI)
```

### Make model dataframes
```{r}
## dung beetle model data, abundance and SR
model_data_db_ab <- drop_na(NeoDungBeetleSites, 
                         RescaledAbundance, Predominant_land_use, Use_intensity, LUI)
model_data_db_sr <- drop_na(NeoDungBeetleSites, 
                         SpeciesRichness, Predominant_land_use, Use_intensity, LUI)

### create alternative abundance metrics

#dung beetle log and sqrt abundances
model_data_db_ab <- mutate(model_data_db_ab, 
                        logAbundance = log(RescaledAbundance + 1),
                        sqrtAbundance = sqrt(RescaledAbundance))
```

### Reorder the model data into the same levels as the abundance data
This may not be a necessary step - however I noticed when I was initially running the code that sometimes the model data gets messed up - best to keep it in any ways to check
```{r}
### DUNG BEETLES

# ABUNDANCE
#reorder levels for abundance LUI
model_data_db_ab$LUI <- factor(model_data_db_ab$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for predominant land use 
model_data_db_ab$Predominant_land_use <- factor(model_data_db_ab$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))


# SPECIES RICHNESS
#reorder levels for Species richness LUI
model_data_db_sr$LUI <- factor(model_data_db_sr$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#predominant land use
model_data_db_sr$Predominant_land_use <- factor(model_data_db_sr$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))
```

### Write the model dataframes as csv so I can easily load it
```{r}
write.csv(model_data_db_ab, "DB_ab_model_data.csv")
```


# MODELLING 

## ABUNDANCE
Now I can move on to modelling the more fine LUI for dung beetles. I will start with the abundance models.

### Create maximal model for dung beetle abundance vs LUI and begin with backwards stepwise selection
```{r}
## Sqrt abundance models -make maximal model
m1 <- lmer(sqrtAbundance ~ LUI + (1|SS) + (1|SSB), data = model_data_db_ab)

## Remove block as random effect
m2 <- lmer(sqrtAbundance ~ LUI + (1|SS), data = model_data_db_ab)

## remove study as random effect
m3 <- lmer(sqrtAbundance ~ LUI + (1|SSB), data = model_data_db_ab)

## Check the AIC values of these models to see the best random effects structure
AIC(m1,m2,m3)

## Model m1 has the lowest AIC and is therefore the random effects structure is the best.

## Proceed with removing the fixed effects - create a null model
m4 <- lmer(sqrtAbundance ~ (1|SS) + (1|SSB), data = model_data_db_ab)

## Compare the models
anova(m1, m4)
model.sel(m1, m4)
AIC(m1, m4)

## Model m1 is better, so that fixed effect structure is the best

## test the LUI model against a model of land use 
m5 <- lmer(sqrtAbundance ~ Predominant_land_use + (1|SS) + (1|SSB), data = model_data_db_ab)

## compare the models
anova(m1, m5)
anova(m5)
Anova(m5)
Anova(m1)
## So LUI vs predominant land use are not significantly different from one another. I will keep LUI as that is the variable I am trying to look at.
```

### Create log abundance model for dung beetle vs LUI, proceed with backwards stepwise selection
```{r}
## Sqrt abundance models -make maximal model
m1a <- lmer(logAbundance ~ LUI + (1|SS) + (1|SSB), data = model_data_db_ab)

## Remove block as random effect
m2a <- lmer(logAbundance ~ LUI + (1|SS), data = model_data_db_ab)

## remove study as random effect
m3a <- lmer(logAbundance ~ LUI + (1|SSB), data = model_data_db_ab)

## Check the AIC values of these models to see the best random effects structure
AIC(m1a,m2a,m3a)

## Model m1a has the lowest AIC and is therefore the random effects structure is the best.

## Proceed with removing the fixed effects - create a null model
m4a <- lmer(logAbundance ~ (1|SS) + (1|SSB), data = model_data_db_ab)

## Compare the models
anova(m1a, m4a)
model.sel(m1a, m4a)
AIC(m1a, m4a)

## Model m1a is better, so that fixed effect structure is the best

## test the LUI model against a model of land use 
m5a <- lmer(logAbundance ~ Predominant_land_use + (1|SS) + (1|SSB), data = model_data_db_ab)

## compare the models
anova(m1a, m5a)

## So LUI vs predominant land use are not significantly different from one another. I will keep LUI as that is the variable I am trying to look at.
```

Models m1 (sqrt) and m1a (log) were significantly better than the others. Time for model validation

### Check model diagnostics for model m4 (sqrt abundance)
```{r}
## Check the residuals using the model_plot function
model_plot(m1)

## Simulate the residuals of the model using DHARMa
m1_residuals <- simulateResiduals(fittedModel = m1, plot = F)

## Plot them
plot(m1_residuals)

## Test the outliers and test for dispersion
testOutliers(m1_residuals)
testDispersion(m1_residuals)
```

### Check model diagnostics for model m1a (log abundance)
```{r}
## Check the residuals using the model_plot function
model_plot(m1a)

## Simulate the residuals of the model using DHARMa
m1a_residuals <- simulateResiduals(fittedModel = m1a, plot = F)

## Plot them
plot(m1a_residuals)

## Test the outliers and test for dispersion
testOutliers(m1a_residuals)
testDispersion(m1a_residuals)
```

The sqrt abundance model residual plots look better than the log + 1 model. So I will probably use that 
Now time to move on to species richness.

### get summary of sqrt abundance model
```{r}
summary(m1)
Anova(m1)
```


## SPECIES RICHNESS

### Create the maximal model and proceed with backwards stepwise selection
```{r}
## Create the maximal model
m1s <- glmer(SpeciesRichness ~ LUI + (1|SS) + (1|SSB), 
            data = model_data_db_sr,
            family = "poisson")

## Remove block as random effect
m2s <- glmer(SpeciesRichness ~ LUI + (1|SS), 
            data = model_data_db_sr,
            family = "poisson")

## Remove study as random effect
m3s <- glmer(SpeciesRichness ~ LUI +(1|SSB), 
            data = model_data_db_sr,
            family = "poisson",
            control = glmerControl(optimizer = "bobyqa"))

## Compare the random effects structure of the model
AIC(m1s, m2s, m3s)
anova(m1s, m2s, m3s)

## Model m1s is the best random effect structure, so keep that.

## Compare the first model to an intercept only model of Species richness
m4s <- glmer(SpeciesRichness ~ (1|SS) + (1|SSB), 
            data = model_data_db_sr,
            family = "poisson")

## Check which model fits better
anova(m1s, m4s)
model.sel(m1s, m4s)

## It seems that model m1s is the best model.
summary(m1s)
```

Time for model validation. 

### Validate the model, check diagnostics of model m1s
```{r}
## run the model function
model_plot(m1s)

## Simulate the residuals of the model using DHARMa
m1s_residuals <- simulateResiduals(fittedModel = m4s, plot = F)

## Plot them
plot(m1s_residuals)

## Test the outliers and test for dispersion
testOutliers(m1s_residuals)
testDispersion(m1s_residuals)

## Test for overdispersion
overdisp_fun(m1s)
```

The overdispersion test for this model was significant, so there is overdispersion in the model. I will add a site level random effect (SSBS) to account for this overdispersion

### Create a model with SSBS as an extra random effect
```{r}
## Add SSBS as random effect
m5s <- glmer(SpeciesRichness ~ LUI + (1|SS) + (1|SSB) + (1|SSBS), 
            data = model_data_db_sr,
            family = "poisson")

## Check the diagnostic plots for this model
model_plot(m5s)

## Simulate the residuals of the model using DHARMa
m5s_residuals <- simulateResiduals(fittedModel = m4s, plot = F)

## Plot them
plot(m5s_residuals)

## Test the outliers and test for dispersion
testOutliers(m5s_residuals)
testDispersion(m5s_residuals)

## Test for overdispersion
overdisp_fun(m5s)
```

Adding in SSBS as an extra random effect fixed the overdispersion. Model m5s will be the species richness model I use to make plots

### check the model summary
```{r}
summary(m5s)
Anova(m5s)
```


Now onto visualizing the model.


# ABUNDANCE MODEL VISUALIZATION

Simulate the effects of the dung beetle abundance models to generate figures

### Simulate the sqrt abundance model
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_ab1_eff <- FEsim(m1, 1000)

# make the term column a factor so it can be recoded
db_ab1_eff$term <- as.factor(db_ab1_eff$term)

# recode the factors to make them shorter (so I can visualise easier)
db_ab1_eff$term <- recode_factor(db_ab1_eff$term, "(Intercept)" = "PVMU",
                              "LUIPVHU" = "PVHU",   
                              "LUIMSV" = "MSV", 
                              "LUIISV" = "ISV", 
                              "LUIYSV" = "YSV", 
                              "LUIPF" = "PF", 
                              "LUIPMU" = "PMU", 
                              "LUIPLU" = "PLU", 
                              "LUIPIU" = "PIU",
                              "LUICMU" = "CMU",
                              "LUICHU" = "CHU")

# Add in upper and lower confidence intervals 
db_ab1_eff <- db_ab1_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the sqrt abundance estimates so that I get a measure of abundance
db_ab1_eff <- db_ab1_eff %>%
  mutate(
    Percent_diff = ((((median[1] + median)^2) / (median[1])^2)*100)-100) %>%
  mutate(
    Percent_upper = (((median[1] + Upper_ci)^2) / ((median[1])^2) * 100) - 100) %>%
  mutate(
    Percent_lower = ((((median[1] + Lower_ci)^2) / ((median[1])^2) * 100) - 100)
  )

# Shift the baseline down to 0
db_ab1_eff[1,7] <- 0 #shift median
db_ab1_eff[1,8] <- 0 #shift upper
db_ab1_eff[1,9] <- 0 #shift lower 

# Rename the sim eff term column name to LUI for joining
colnames(db_ab1_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_data_db_ab_plot <- left_join(model_data_db_ab, db_ab1_eff)
```

### Plot the effect estimates 
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
DB_sqrt_abundance_plot <- model_data_db_ab_plot %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, colour = Predominant_land_use)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = Percent_lower, ymax = Percent_upper), size = 0.75, width = 0.2)+
  geom_hline(yintercept = 0, size = I(1), color = I("black"), linetype = 1 )+
  theme_light()+
  theme(axis.text.x = element_text(colour = c("#000000", "#000000", "#E69F00", "#56B4E9", "#009E73",  "#F0E442", "#0072B2", "#0072B2", "#0072B2","#D55E00", "#D55E00"), face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank())+
  ylab("Total abundance difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  theme(legend.position = "none")
DB_sqrt_abundance_plot
```
Figure X: Dung beetle sqrt abundance response to land use intensity (LUI) in the Neotropics
Error bars show the 95% confidence interval around each estimate. Error bars that do not cross zero are statistically significantly different from primary vegetation minimal use. Colors represent the following land use types: black (primary vegetation, PV), orange (mature secondary vegetation, MSV), light blue (intermediate secondary vegetation, ISV), green (young secondary vegetation, YSV), yellow (plantation forest, PF), dark blue (pasture), dark orange (cropland). N = 995 observations

### Simulate the log abundance model
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_ab2_eff <- FEsim(m1a, 1000)

# make the term column a factor so it can be recoded
db_ab2_eff$term <- as.factor(db_ab2_eff$term)

# recode the factors to make them shorter (so I can visualise easier)
db_ab2_eff$term <- recode_factor(db_ab2_eff$term, "(Intercept)" = "PVMU",
                              "LUIPVHU" = "PVHU",   
                              "LUIMSV" = "MSV", 
                              "LUIISV" = "ISV", 
                              "LUIYSV" = "YSV", 
                              "LUIPF" = "PF", 
                              "LUIPMU" = "PMU", 
                              "LUIPLU" = "PLU", 
                              "LUIPIU" = "PIU",
                              "LUICMU" = "CMU",
                              "LUICHU" = "CHU")

# Add in upper and lower confidence intervals 
db_ab2_eff <- db_ab2_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the log abundance estimates so that I get a measure of abundance
db_ab2_eff <- db_ab2_eff %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

# Shift the baseline down to 0
db_ab2_eff[1,7] <- 0 #shift median
db_ab2_eff[1,8] <- 0 #shift upper
db_ab2_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(db_ab2_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_data_db_ab_plot2 <- left_join(model_data_db_ab, db_ab2_eff)
```

### Plot the effect estimates 
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
DB_log_abundance_plot <- model_data_db_ab_plot2 %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, colour = Predominant_land_use)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = Percent_lower, ymax = Percent_upper), size = 0.75, width = 0.2)+
  geom_hline(yintercept = 0, size = I(1), color = I("black"), linetype = 1 )+
  theme_light()+
  theme(axis.text.x = element_text(colour = c("#000000", "#000000", "#E69F00", "#56B4E9", "#009E73",  "#F0E442", "#0072B2", "#0072B2", "#0072B2","#D55E00", "#D55E00"), face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank())+
  xlab("Land use intensity class")+
  ylab("Total abundance difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  theme(legend.position = "none")
DB_log_abundance_plot
```
Figure X: (LOG) Dung beetle total abundance response to land use intensity (LUI) in the Neotropics.
Error bars show the 95% confidence interval around each estimate. Error bars that do not cross zero are statistically significantly different from primary vegetation minimal use. Colors represent the following land use types: black (primary vegetation, PV), orange (mature secondary vegetation, MSV), light blue (intermediate secondary vegetation, ISV), green (young secondary vegetation, YSV), yellow (plantation forest, PF), dark blue (pasture), dark orange (cropland). N = 995 observations

# SPECIES RICHNESS VISUALIZATION

### Simulate the dung beetle species richness model
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_sr_eff <- FEsim(m5s, 1000)

# make the term column a factor so it can be recoded
db_sr_eff$term <- as.factor(db_sr_eff$term)

# recode the factors to make them shorter (so I can visualise easier)
db_sr_eff$term <- recode_factor(db_sr_eff$term, "(Intercept)" = "PVMU",
                              "LUIPVHU" = "PVHU",   
                              "LUIMSV" = "MSV", 
                              "LUIISV" = "ISV", 
                              "LUIYSV" = "YSV", 
                              "LUIPF" = "PF", 
                              "LUIPMU" = "PMU", 
                              "LUIPLU" = "PLU", 
                              "LUIPIU" = "PIU",
                              "LUICMU" = "CMU",
                              "LUICHU" = "CHU")

# Add in upper and lower confidence intervals 
db_sr_eff <- db_sr_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Generate percentage differences from the baseline (first term, Primary Vegetation (PV))
db_sr_eff <- db_sr_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
db_sr_eff[1,7] <- 0 #shift median
db_sr_eff[1,8] <- 0 #shift upper
db_sr_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(db_sr_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_data_db_sr_plot <- left_join(model_data_db_sr, db_sr_eff)
```

### Plot the effect estimates 
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
DB_sr_plot <- model_data_db_sr_plot %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, colour = Predominant_land_use)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = Percent_lower, ymax = Percent_upper), size = 0.75, width = 0.2)+
  geom_hline(yintercept = 0, size = I(1), color = I("black"), linetype = 1 )+
  theme_light()+
  theme(axis.text.x = element_text(colour = c("#000000", "#000000", "#E69F00", "#56B4E9", "#009E73",  "#F0E442", "#0072B2", "#0072B2", "#0072B2","#D55E00", "#D55E00"), face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank())+
  xlab("Land use intensity class")+
  ylab("Species richness difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  theme(legend.position = "none")
DB_sr_plot
```
Figure X: Dung beetle species richness response to land use intensity (LUI) in the Neotropics
Error bars show the 95% confidence interval around each estimate. Error bars that do not cross zero are statistically significantly different from primary vegetation minimal use. Colors represent the following land use types: black (primary vegetation, PV), orange (mature secondary vegetation, MSV), light blue (intermediate secondary vegetation, ISV), green (young secondary vegetation, YSV), yellow (plantation forest, PF), dark blue (pasture), dark orange (cropland). N = 995 observations


### Save the model dataframes with which I made the graphs for a separate plotting script
```{r}
## Sqrt abundance
write.csv(model_data_db_ab_plot, "DB_abundance_data_for_plots.csv")

## Species richness
write.csv(model_data_db_sr_plot, "DB_species_richness_data_for_plots.csv")
```

I have saved the model data separately as I want to combine all the graphs together - I also might want to change how they look, so it is easier to do it this way.