---
title: "3a_DB_Simpsons_Index"
author: "Alex Dhond"
date: "8/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for calculating Simpson's diversity index for the dung beetle sites in my dataset. The actual modelling of Simpson's diversity comes in the next script (script 3b)

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(vegan) # for calculating Simpson's index
```

### First read in the Dung beetle data
```{r}
NeoDungBeetles <- read.csv("NeoDungBeetles.csv")
```

### For Simpson's diversity index, I need a measure of the abundance of unique dung beetle species at each site
```{r}
## Remove rows where there is no genus information (this should have been done already in the first script, but just to check again)
NeoDungBeetleSpecies <- droplevels(subset(NeoDungBeetles, Genus != ""))

# Create a new GenusSpecies column, where if the species name is unknown, it will put in the genus name instead.
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%
  mutate(GenusSpecies = ifelse(Best_guess_binomial == "", paste(Genus), paste(Best_guess_binomial)))
```

### Generate the abundance of each species at each site
```{r}
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%

  # pull out only the merged diversity data
  distinct(merge_ID, .keep_all = TRUE) %>%
  
  # re-make SSB and SSBS values since we've now dropped a bunch of values
  mutate(SS = paste(Source_ID, Study_number),
         SSB = paste(SS, Block),
         SSBS = paste(SSB, Site_number)) %>%
  
  # group by SSBS (each unique value corresponds to a unique site)
  group_by(SSBS, GenusSpecies) %>%
  
  # now add up all the abundance measurements within each site
  mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                 sum(merged_diversity),
                                 # if the diversity metric type isn't Abundance, then leave the TotalAbundance measurement as NA
                                 NA),
         
         # if the metric is already species richness
         SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                  # just use the value as given
                                  merged_diversity,
                                  # for abundance and occurrence measurements, count the number of unique species names that are present at the site 
                                  n_distinct(Taxon_name_entered[merged_diversity > 0])),
         
         # calculate Chao's Species Richness Estimator
         # if the diversity metric is suitable for this calculation
         ChaoRichness = ifelse(Diversity_metric_is_suitable_for_Chao == TRUE,
                               # calculate the Chao estimator
                               sum(merged_diversity > 0) + (((sum(merged_diversity == 1) * (sum(merged_diversity == 1)-1)) / (2*(sum(merged_diversity == 2)+1)))),
                               # otherwise give Chao NA
                               NA)
         ) %>%
  
  # ungroup
  ungroup() %>%
  
    # now group by Study ID
  group_by(SS, GenusSpecies) %>%
  
  # pull out some useful study-level numbers
  # maximum abundance for each study
  mutate(MaxAbundance = max(TotalAbundance),
         # number of species in the study
         SpeciesInStudy = n_distinct(Taxon_name_entered[merged_diversity > 0]),
         # assess whether the study is suitable for rarefaction
         SuitableForRarefaction = ifelse(
           # if all diversity measurements are integers
           # i.e. if you round down, it should be equal to the original number
           # and the diversity metric is suitable for Chao
           all(floor(merged_diversity) == merged_diversity) &
             Diversity_metric_is_suitable_for_Chao == TRUE,
           # then class the study as suitable for rarefaction
           TRUE,
           # otherwise it can't be used
           FALSE
           )
         ) %>%
  
  # ungroup
  ungroup()  %>%
  
  # now rescale total abundance, so that within each study, abundance varies from 0 to 1.
  mutate(RescaledAbundance = TotalAbundance/MaxAbundance,
         # for statistical modelling, we'll also calculate the square root of species abundance, although we might want to use log(x+1) transformation instead
         sqrtRescaledAbundance = sqrt(RescaledAbundance),
         logRescaledAbundance = log(RescaledAbundance + 1)
         )
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%
  distinct(SSBS, GenusSpecies, .keep_all = TRUE)
```


Within the NeoDungBeetleSpecies dataframe, the "TotalAbundance" column gives the total abundance of that particular species at that site. The "SpeciesRichness" column gives either a 0 or 1 depending on if that species is present or absent - this is useful in terms of generating a presence absence dataframe

I now need to collapse the land use intensities in the same way that I did the dung beetle sites data (script 3) so that I can compare Simpson's to species richness

### Look at the site data, and collapse variables with little/no data
```{r}
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%
  
  mutate(

    #collapse primary vegetation together
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                  "Primary forest" = "Primary vegetation", 
                                  "Primary non-forest" = "Primary vegetation"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Use_intensity = na_if(Use_intensity, "Cannot decide"),

  )

#remove urban sites
NeoDungBeetleSpecies <- droplevels(subset(NeoDungBeetleSpecies, Predominant_land_use != "Urban"))

# take another look at the LandUse/Use intensity split
table(NeoDungBeetleSpecies$Predominant_land_use, NeoDungBeetleSpecies$Use_intensity) 
```

### Collapse the land use categories further
```{r}
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%
  mutate(LUI = interaction(Predominant_land_use, Use_intensity, sep = "_"),
         
         #collapse primary vegetation Light Intense together
         LUI = recode_factor(LUI, "Primary vegetation_Light use" = "Primary vegetation_LightIntense", "Primary vegetation_Intense use" = "Primary vegetation_LightIntense"),
         
         #collapse Cropland Intense and Light together, save as "LightIntense"
         LUI = recode_factor(LUI, "Cropland_Intense use" = "Cropland_LightIntense",
                             "Cropland_Light use" = "Cropland_LightIntense"),
         
         #collapse Mature Secondary Vegetation Intense and Light into "LightIntense"
         LUI = recode_factor(LUI, 
                             "Mature secondary vegetation_Minimal use" = "Mature secondary vegetation",
                             "Mature secondary vegetation_Intense use" = "Mature secondary vegetation", 
                             "Mature secondary vegetation_Light use" = "Mature secondary vegetation"),
         
         #collapse plantation forest intense and light together
         LUI = recode_factor(LUI, "Plantation forest_Intense use" = "Plantation forest",
                             "Plantation forest_Light use" = "Plantation forest",
                             "Plantation forest_Minimal use" = "Plantation forest"),
         
         #collapse Intermediate secondary vegetation light and intense together
         LUI = recode_factor(LUI, "Intermediate secondary vegetation_Intense use" = "Intermediate secondary vegetation", "Intermediate secondary vegetation_Light use" = "Intermediate secondary vegetation",
                             "Intermediate secondary vegetation_Minimal use" = "Intermediate secondary vegetation"),
         
         #collapse Young secondary vegetation light and intense together
         LUI = recode_factor(LUI, "Young secondary vegetation_Light use" = "Young secondary vegetation", "Young secondary vegetation_Intense use" = "Young secondary vegetation",
                             "Young secondary vegetation_Minimal use" = "Young secondary vegetation"),

         #dont collapse pasture
  )

table(NeoDungBeetleSpecies$LUI)
table(NeoDungBeetleSpecies$Predominant_land_use, NeoDungBeetleSpecies$Use_intensity)
```

### reorder the LUI factor levels for easier visualization
```{r}
#LightIntense use category will be classified as "Higher (H)" use for brevity
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%
  mutate(LUI = plyr::revalue(LUI, c("Primary vegetation_Minimal use" = "PVMU",
                                    "Primary vegetation_LightIntense" = "PVHU",
                                    "Mature secondary vegetation" = "MSV",
                                    "Intermediate secondary vegetation" = "ISV",
                                    "Young secondary vegetation" = "YSV",
                                    "Plantation forest" = "PF",
                                    "Pasture_Minimal use" = "PMU",
                                    "Pasture_Light use" = "PLU",
                                    "Pasture_Intense use" = "PIU",
                                    "Cropland_Minimal use" = "CMU",
                                    "Cropland_LightIntense" = "CHU")))
```

### reorder LUI factor levels
```{r}
#first get a list of the levels
levels(NeoDungBeetleSpecies$LUI)

#recode the levels
NeoDungBeetleSpecies$LUI <- factor(NeoDungBeetleSpecies$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#check it worked
levels(NeoDungBeetleSpecies$LUI)
```

### Do the same for the predominant land use categories
First revalue/relabel the land use categories so that graphs are easier to read. Then recode the levels so that when modelling they appear with primary vegetation as the reference level (and on graphs they go in an intuitive order)
```{r}
## Revalue the factors
NeoDungBeetleSpecies <- NeoDungBeetleSpecies %>%
  mutate(Predominant_land_use = plyr::revalue(Predominant_land_use, c("Primary vegetation" = "Primary",
                                    "Mature secondary vegetation" = "MSV",
                                    "Intermediate secondary vegetation" = "ISV",
                                    "Young secondary vegetation" = "YSV",
                                    "Plantation forest" = "Plantation",
                                    "Pasture" = "Pasture",
                                    "Cropland" = "Cropland")))

## Check to see the names are new
table(NeoDungBeetleSpecies$Predominant_land_use)

## Now re-code the levels
levels(NeoDungBeetleSpecies$Predominant_land_use)

#recode the levels
NeoDungBeetleSpecies$Predominant_land_use <- factor(NeoDungBeetleSpecies$Predominant_land_use, levels = c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland"))

#check it worked
levels(NeoDungBeetleSpecies$Predominant_land_use)
```

### make model data and visualise
```{r}
#dung beetle model data, abundance and SR
model_data_simp_db_ab <- drop_na(NeoDungBeetleSpecies, 
                         RescaledAbundance, Predominant_land_use, Use_intensity, LUI)

#dung beetle log and sqrt abundances
model_data_simp_db_ab <- mutate(model_data_simp_db_ab, 
                        logAbundance = log(RescaledAbundance + 1),
                        sqrtAbundance = sqrt(RescaledAbundance))

# species richness
model_data_simp_db_sr <- drop_na(NeoDungBeetleSpecies, 
                         SpeciesRichness, Predominant_land_use, Use_intensity, LUI)
```

### reorder the model data into the same levels as the abundance data
```{r}
### DUNG BEETLES

# ABUNDANCE
#reorder levels for abundance LUI
model_data_simp_db_ab$LUI <- factor(model_data_simp_db_ab$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for abundance Predominant land use
model_data_simp_db_ab$Predominant_land_use <- factor(model_data_simp_db_ab$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))

#reorder levels for abundance Use intensity
model_data_simp_db_ab$Use_intensity <- factor(model_data_simp_db_ab$Use_intensity, levels = c("Minimal use", "Light use", "Intense use"))

# SPECIES RICHNESS

#reorder levels for Species richness LUI
model_data_simp_db_sr$LUI <- factor(model_data_simp_db_sr$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for Species richness Predominant land use
model_data_simp_db_sr$Predominant_land_use <- factor(model_data_simp_db_sr$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))

#reorder levels for Species Richness Use intensity
model_data_simp_db_sr$Use_intensity <- factor(model_data_simp_db_sr$Use_intensity, levels = c("Minimal use", "Light use", "Intense use"))
```

Now I should be able to calculate Simpson's index (normal and inverse) for all of the sites in my Dung beetles dataset. Here I will calculate both indices for each site, then left join them back into the larger dataset so that each site gets a measure of Simpson's diversity.

## Calculate Simpson's diversity index for dung beetles
```{r}
## create initial dataframe of all the species and their abundances at the site
Simpson_df <- model_data_simp_db_ab %>%
  dplyr::select(SSBS, GenusSpecies, TotalAbundance)

## make it wide data so each species is its own column
Simpson_matrix <- Simpson_df %>%
  pivot_wider(names_from = GenusSpecies, values_from = TotalAbundance)

## remove NAs
Simpson_matrix[is.na(Simpson_matrix)] <- 0

## make site a factor
Simpson_matrix$SSBS <- as.factor(Simpson_matrix$SSBS)

#generate inverse simpsons index
DB_inv_simp <- plyr::ddply(Simpson_matrix, ~SSBS, function(x) {
  data.frame(INVSIMPSON=diversity(x[-1], index="invsimpson"))
})
```

Now I have the values for Simpson's diversity index as well as the Inverse Simpson's diversity index, however in the inverse diversity index there are NAs and Infinite values. I need to remove all the sites where there are no species present as that is what is creating an error in the index.

### Remove all the 0s
```{r}
## Remove all rows where the sum of the abundances is not greater than zero (ie there are no species in that row)
Simp_fixed <- Simpson_matrix[rowSums(Simpson_matrix[2:406])>0,]
995-924
## It removed 71 rows

#generate inverse simpsons index
Simp_inv_fixed_results <- plyr::ddply(Simp_fixed, ~SSBS, function(x) {
  data.frame(INVSIMPSON=diversity(x[-1], index="invsimpson"))
})

## make a log transformed column
Simp_inv_fixed_results$LOGINVSIMPSON <- log(Simp_inv_fixed_results$INVSIMPSON + 1)

```

I now have a dataframe with the Inverse Simpson's diversity index for all of the dung beetle sites in my analysis. I need to now left join this with the model data in order to be able to compare models.

### Save as a csv file for easier loading
```{r}
write.csv(Simp_inv_fixed_results, "DB_Simpson_short.csv")
```


This script calculated inverse simpsons index for the dung beetle sites at which there was at least 1 species sampled.




