---
title: "3b_DB_Simpsons_Models"
author: "Alex Dhond"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for modelling Simpson's diversity for dung beetles at the sites in the Neotropics.

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation

## Modelling ## 
library(lme4) # creating mixed effects models
library(car) # for LRT / anova comparisons
library(merTools) # for extracting model estimates
library(MuMIn) # for assessing model fit
library(optimx) #optimizing models
library(DHARMa) #for model validation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in required functions
```{r}
## Load in the source function for checking GVIFs - for collinearity
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(1,3))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}

## Load in an overdispersion function from this website to help test if the models are overdispersed.
# https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-for-overdispersioncomputing-overdispersion-factor
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

### Load in plotting variables and colors
```{r}
## for the land use classes
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### Read in the Simpson csv file as well as the DB sites data and join them
```{r}
## Read in the simpson file
DB_Simpson <- read.csv("DB_Simpson_short.csv")

## remove the first column so it doesnt join by that
DB_Simpson <- DB_Simpson[,-1]

## Read in the abundance model data
DB_abundance <- read.csv("DB_ab_model_data.csv")

## Create the model data for simpson
DB_simp_model <- left_join(DB_abundance, DB_Simpson)
```

### Drop all the NAs from the model and reorder factor levels
```{r}
DB_simp_model <- drop_na(DB_simp_model,INVSIMPSON, Predominant_land_use, LUI)

#reorder levels for LUI
DB_simp_model$LUI <- factor(DB_simp_model$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for Predominant land use
DB_simp_model$Predominant_land_use <- factor(DB_simp_model$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))
```

Now that the data is ready, make some models of inverse simpsons index
### Make the maximal model and proceed with backwards stepwise selection
```{r}
## Create the maximal model
m1 <- lmer(INVSIMPSON ~ LUI + (1|SS) + (1|SSB), data = DB_simp_model)

## Remove block as random effect
m2 <- lmer(INVSIMPSON ~ LUI + (1|SS), data = DB_simp_model)

## Remove study as random effect
m3 <- lmer(INVSIMPSON ~ LUI + (1|SSB), data = DB_simp_model)

## Compare the AIC to determine best random effects structure
AIC(m1,m2,m3)

## Model m1 is the lowest AIC, so keep that random effects structure

## Now test fixed effects

## Create an intercept only model
m4 <- lmer(INVSIMPSON ~ (1|SS) + (1|SSB), data = DB_simp_model)

## Compare the maximal model to the null model
AIC(m1, m4)
anova(m1,m4)
model.sel(m1,m4)

## Model m1 is significantly better, so check model diagnostic plots
```

### Check the model diagnostic plots
```{r}
# look at the model criticism plots
model_plot(m1)
## Simulate the residuals of the model using DHARMa
m1_residuals <- simulateResiduals(fittedModel = m1, plot = F)

## Plot them
plot(m1_residuals)

## Test the outliers and test for dispersion
testOutliers(m1_residuals)
testDispersion(m1_residuals)
```

There seems to be a bit of skew, so I am now going to run a model of Log + 1 transformed inverse simpsons index

### Create a model for the log+1 transformed Simpsons index and check diagnostics
```{r}
## Create model
m5 <- lmer(LOGINVSIMPSON ~ LUI + (1|SS) + (1|SSB), data = DB_simp_model)

# look at the model criticism plots
model_plot(m5)
## Simulate the residuals of the model using DHARMa
m5_residuals <- simulateResiduals(fittedModel = m1, plot = F)

## Plot them
plot(m5_residuals)

## Test the outliers and test for dispersion
testOutliers(m5_residuals)
testDispersion(m5_residuals)
```

Log + 1 transforming the inverse simpsons index looks a lot better than the normal inverse simpsons index, so I will use this model for plotting

### get model estimates
```{r}
summary(m5)
anova(m5)
Anova(m5)
```


### Simulate the Simpsons model effects
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
db_simp_eff <- FEsim(m5, 1000)

# make the term column a factor so it can be recoded
db_simp_eff$term <- as.factor(db_simp_eff$term)

# recode the factors to make them shorter (so I can visualise easier)
db_simp_eff$term <- recode_factor(db_simp_eff$term, "(Intercept)" = "PVMU",
                              "LUIPVHU" = "PVHU",   
                              "LUIMSV" = "MSV", 
                              "LUIISV" = "ISV", 
                              "LUIYSV" = "YSV", 
                              "LUIPF" = "PF", 
                              "LUIPMU" = "PMU", 
                              "LUIPLU" = "PLU", 
                              "LUIPIU" = "PIU",
                              "LUICMU" = "CMU",
                              "LUICHU" = "CHU")

# Add in upper and lower confidence intervals 
db_simp_eff <- db_simp_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the logged estimates so that I get a measure of Simpson's diversity
db_simp_eff <- db_simp_eff %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

# Shift the baseline down to 0
db_simp_eff[1,7] <- 0 #shift median
db_simp_eff[1,8] <- 0 #shift upper
db_simp_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(db_simp_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
Simpson_plot <- left_join(DB_simp_model, db_simp_eff)
```

### Plot the effect estimates 
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
Simpson_plot %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, colour = Predominant_land_use)+
  geom_point(size = 2)+
  geom_errorbar(aes(ymin = Percent_lower, ymax = Percent_upper), size = 0.75, width = 0.2)+
  geom_hline(yintercept = 0, size = I(1), color = I("black"), linetype = 1 )+
  theme_light()+
  theme(axis.text.x = element_text(colour = c("#000000", "#000000", "#E69F00", "#56B4E9", "#009E73",  "#F0E442", "#0072B2", "#0072B2", "#0072B2","#D55E00", "#D55E00"), face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank())+
  xlab("Land use intensity class")+
  ylab("Simpson's diversity difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  theme(legend.position = "none")
Simpson_plot
```
Figure X: Dung beetle Simpson's diversity response to land use intensity (LUI) in the Neotropics.
Error bars show the 95% confidence interval around each estimate. Error bars that do not cross zero are statistically significantly different from primary vegetation minimal use. Colors represent the following land use types: black (primary vegetation, PV), orange (mature secondary vegetation, MSV), light blue (intermediate secondary vegetation, ISV), green (young secondary vegetation, YSV), yellow (plantation forest, PF), dark blue (pasture), dark orange (cropland). N = 924 observations

### Save the model data so I can plot it
```{r}
## Save the model data as a CSV
write.csv(Simpson_plot, "DB_Simpson_plotting.csv")
```