---
title: "3c_DB_Diversity_Plots"
author: "Alex Dhond"
date: "8/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is for loading in the dung beetle biodiversity data, plotting the figures, combining them into one figure, and then exporting to PDF

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in any colors and variables
```{r}
## Load in a color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### Load in the figure files
```{r}
## Load in abundance plot
DB_abundance_plot <- read.csv("DB_abundance_data_for_plots.csv")

## reorder levels for lui
DB_abundance_plot$LUI <- factor(DB_abundance_plot$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for Predominant land use
DB_abundance_plot$Predominant_land_use <- factor(DB_abundance_plot$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))


## Load in species richness plot
DB_speciesrichness_plot <- read.csv("DB_species_richness_data_for_plots.csv")

## reorder levels for lui
DB_speciesrichness_plot$LUI <- factor(DB_speciesrichness_plot$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for Predominant land use
DB_speciesrichness_plot$Predominant_land_use <- factor(DB_speciesrichness_plot$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))


## Load in simpson plot
DB_simpson_plot <- read.csv("DB_Simpson_plotting.csv")
#reorder levels for LUI
DB_simpson_plot$LUI <- factor(DB_simpson_plot$LUI, levels = c("PVMU", "PVHU", "MSV", "ISV", "YSV", "PF", "PMU", "PLU", "PIU", "CMU", "CHU"))

#reorder levels for Predominant land use
DB_simpson_plot$Predominant_land_use <- factor(DB_simpson_plot$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))
```

### For plotting, make a new "use intensity" SPECIES RICHNESS
```{r}
## make a new use intensity column
DB_speciesrichness_plot <- DB_speciesrichness_plot %>%
  mutate(
    Use_intensity2 = ifelse(LUI == "PVMU" | LUI == "CMU" | LUI == "PMU", "Minimal use", ifelse(LUI == "PVHU" | LUI == "CHU", "Higher use", ifelse(LUI == "PLU", "Light use", ifelse(LUI == "PIU", "Intense use", "Average"))))
  )

## make it a factor
DB_speciesrichness_plot$Use_intensity2 <- factor(DB_speciesrichness_plot$Use_intensity2, levels = c("Average", "Minimal use", "Light use", "Higher use", "Intense use"))

levels(DB_speciesrichness_plot$Use_intensity2)
```

### For plotting, make a new "use intensity" ABUNDANCE
```{r}
## make a new use intensity column
DB_abundance_plot <- DB_abundance_plot %>%
  mutate(
    Use_intensity2 = ifelse(LUI == "PVMU" | LUI == "CMU" | LUI == "PMU", "Minimal use", ifelse(LUI == "PVHU" | LUI == "CHU", "Higher use", ifelse(LUI == "PLU", "Light use", ifelse(LUI == "PIU", "Intense use", "Average"))))
  )

## make it a factor
DB_abundance_plot$Use_intensity2 <- factor(DB_abundance_plot$Use_intensity2, levels = c("Average", "Minimal use", "Light use", "Higher use", "Intense use"))

levels(DB_abundance_plot$Use_intensity2)
```

### For plotting, make a new "use intensity" SIMPSONS
```{r}
## make a new use intensity column
DB_simpson_plot <- DB_simpson_plot %>%
  mutate(
    Use_intensity2 = ifelse(LUI == "PVMU" | LUI == "CMU" | LUI == "PMU", "Minimal use", ifelse(LUI == "PVHU" | LUI == "CHU", "Higher use", ifelse(LUI == "PLU", "Light use", ifelse(LUI == "PIU", "Intense use", "Average"))))
  )

## make it a factor
DB_simpson_plot$Use_intensity2 <- factor(DB_simpson_plot$Use_intensity2, levels = c("Average", "Minimal use", "Light use", "Higher use", "Intense use"))

levels(DB_simpson_plot$Use_intensity2)
```

### Set up all things needed for plots
```{r}
## set up the limits along the x axis
DB_X_Limits <- c("PVMU", "PVHU",
              "MSV",
              "ISV",
              "YSV",
              "PF",
              "PMU",
              "PLU",
              "PIU",
              "CMU",
              "CHU")

## set up land use limits 
DB_X_Limits_2 <- c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland")
```

### Make the plots 
```{r}
## Make species richness plot
DB_sr_figure <- DB_speciesrichness_plot %>%
  ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper, colour = Predominant_land_use, shape = Use_intensity2)+
  geom_hline(yintercept = 0, size = 0.5, color = "black", linetype = 1)+
  geom_point(aes(shape = Use_intensity2), size = 4, position = position_dodge(width = 0.75))+
  geom_linerange(size = 1, position = position_dodge(width = 0.75))+
  scale_x_discrete(limits=DB_X_Limits_2, labels=c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland")) + xlab("") + ## can add labels = c("")..
  theme_classic()+
  theme(axis.text.x = element_text(colour = cbbPalette, face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Species richness difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  scale_shape_manual(values=c(4, 16, 17, 18, 15))+
  theme(legend.position = "none")
DB_sr_figure

## Make abundance plot
DB_abundance_figure <- DB_abundance_plot %>%
ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper, colour = Predominant_land_use, shape = Use_intensity2)+
  geom_hline(yintercept = 0, size = 0.5, color = "black", linetype = 1)+
  geom_point(aes(shape = Use_intensity2), size = 4, position = position_dodge(width = 0.75))+
  geom_linerange(size = 1, position = position_dodge(width = 0.75))+
  scale_x_discrete(limits=DB_X_Limits_2, labels=c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland")) + xlab("") + ## can add labels = c("")..
  theme_classic()+
  theme(axis.text.x = element_text(colour = cbbPalette, face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Total abundance difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  scale_shape_manual(values=c(4, 16, 17, 18, 15))+
  theme(legend.position = "none")
DB_abundance_figure

## Make simpsons diversity plot 
DB_simpson_figure <- DB_simpson_plot %>%
ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper, colour = Predominant_land_use, shape = Use_intensity2)+
  geom_hline(yintercept = 0, size = 0.5, color = "black", linetype = 1)+
  geom_point(aes(shape = Use_intensity2), size = 4, position = position_dodge(width = 0.75))+
  geom_linerange(size = 1, position = position_dodge(width = 0.75))+
  scale_x_discrete(limits=DB_X_Limits_2, labels=c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland")) + xlab("") + ## can add labels = c("")..
  theme_classic()+
  theme(axis.text.x = element_text(colour = cbbPalette, face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Simpson's diversity difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  scale_shape_manual(values=c(4, 16, 17, 18, 15))+
  theme(legend.position = "none")
DB_simpson_figure
```

### make the legend for the plot
```{r}
## create a new plot, extract the legend
DB_sr_figure_legend <- DB_speciesrichness_plot %>%
  ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper, colour = Predominant_land_use, shape = Use_intensity2)+
  geom_hline(yintercept = 0, size = 0.5, color = "black", linetype = 1)+
  geom_point(aes(shape = Use_intensity2), size = 4, position = position_dodge(width = 0.75))+
  geom_linerange(size = 1, position = position_dodge(width = 0.75))+
  scale_x_discrete(limits=DB_X_Limits_2, labels=c("Primary", "MSV", "ISV", "YSV", "Plantation", "Pasture", "Cropland")) + xlab("") + ## can add labels = c("")..
  theme_classic()+
  theme(axis.text.x = element_text(colour = cbbPalette, face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Species richness difference (%)")+
  scale_colour_manual(values=cbbPalette)+
  scale_shape_manual(values=c(4, 16, 17, 18, 15))+
  theme(legend.position = "bottom",
        legend.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"))+
  guides(shape = guide_legend("Use Intensity:"),
         color = F)

# Using the cowplot package
DB_legend <- cowplot::get_legend(DB_sr_figure_legend)
DB_legend
DB_plot_legend <- as_ggplot(DB_legend)
DB_plot_legend
```

### Combine all the plots together and export
```{r}
DB_diversity_plots_1 <- ggarrange(DB_abundance_figure, DB_sr_figure,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2)
DB_diversity_plots_2 <- ggarrange(DB_simpson_figure, DB_plot_legend,
                    labels = c("C", ""),
                    ncol = 1, nrow = 2)

ggexport(DB_diversity_plots_1, filename = "DB_diversity_plots_1.pdf")
ggexport(DB_diversity_plots_2, filename = "DB_diversity_plots_2.pdf")
```