---
title: "4_Ant_Site_Diversity"
author: "Alex Dhond"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for testing Ant data site metrics. In this script, I want to get a general overview of ant species richness for land-use intensity across my sites in the Neotropics. I hope to see broad trends in abundance and species richness across different land uses and use intensities

Due to the data spread, and the fact that most ant data is occurrence measurements, I am not able to get a good measure of ant abundance. Therefore, I am only calculating species richness in this script.

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation

## Modelling ## 
library(lme4) # creating mixed effects models
library(car) # for LRT / anova comparisons
library(merTools) # for extracting model estimates
library(MuMIn) # for assessing model fit
library(optimx) #optimizing models
library(DHARMa) #for model validation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in any required functions
```{r}
## Load in the source function for checking GVIFs - for collinearity
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(1,3))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}

## Load in an overdispersion function from this website to help test if the models are overdispersed.
# https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-for-overdispersioncomputing-overdispersion-factor
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

### Load in a colour palette (for the figures)
```{r}
## Load in a color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### Read in the Ants data
```{r}
NeoAnts <- read.csv("NeoAnts.csv")
```

### Generate the Neotropical ant sites
```{r}
NeoAntSites <- NeoAnts %>%

  # pull out only the merged diversity data
  distinct(merge_ID, .keep_all = TRUE) %>%
  
  # re-make SSB and SSBS values since we've now dropped a bunch of values
  mutate(SS = paste(Source_ID, Study_number),
         SSB = paste(SS, Block),
         SSBS = paste(SSB, Site_number)) %>%
  
  # group by SSBS (each unique value corresponds to a unique site)
  group_by(SSBS) %>%
  
  # now add up all the abundance measurements within each site
  mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                 sum(merged_diversity),
                                 # if the diversity metric type isn't Abundance, then leave the TotalAbundance measurement as NA
                                 NA),
         
         # if the metric is already species richness
         SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                  # just use the value as given
                                  merged_diversity,
                                  # for abundance and occurrence measurements, count the number of unique species names that are present at the site 
                                  n_distinct(Taxon_name_entered[merged_diversity > 0])),
         
         # calculate Chao's Species Richness Estimator
         # if the diversity metric is suitable for this calculation
         ChaoRichness = ifelse(Diversity_metric_is_suitable_for_Chao == TRUE,
                               # calculate the Chao estimator
                               sum(merged_diversity > 0) + (((sum(merged_diversity == 1) * (sum(merged_diversity == 1)-1)) / (2*(sum(merged_diversity == 2)+1)))),
                               # otherwise give Chao NA
                               NA)
         ) %>%
  
  # ungroup
  ungroup() %>%
  
    # now group by Study ID
  group_by(SS) %>%
  
  # pull out some useful study-level numbers
  # maximum abundance for each study
  mutate(MaxAbundance = max(TotalAbundance),
         # minimum (non-zero) abundance for each study
         # we'll use this when we do species rarefaction
         MinNonZeroAbundance = min(TotalAbundance[TotalAbundance > 0]),
         # number of species in the study
         SpeciesInStudy = n_distinct(Taxon_name_entered[merged_diversity > 0]),
         # assess whether the study is suitable for rarefaction
         SuitableForRarefaction = ifelse(
           # if all diversity measurements are integers
           # i.e. if you round down, it should be equal to the original number
           # and the diversity metric is suitable for Chao
           all(floor(merged_diversity) == merged_diversity) &
             Diversity_metric_is_suitable_for_Chao == TRUE,
           # then class the study as suitable for rarefaction
           TRUE,
           # otherwise it can't be used
           FALSE
           )
         ) %>%
  
  # ungroup
  ungroup()  %>%
  
  # now rescale total abundance, so that within each study, abundance varies from 0 to 1.
  mutate(RescaledAbundance = TotalAbundance/MaxAbundance,
         # for statistical modelling, we'll also calculate the square root of species abundance, although we might want to use log(x+1) transformation instead
         sqrtRescaledAbundance = sqrt(RescaledAbundance),
         logRescaledAbundance = log(RescaledAbundance + 1)
         )

# pull out unique sites
NeoAntSites <- NeoAntSites %>%
  distinct(SSBS, .keep_all = TRUE)
```

### check the number of sources, studies, and sites
```{r}
#ants
length(unique(NeoAntSites$Source_ID)) #number of sources
length(unique(NeoAntSites$SS)) #number of studies
length(unique(NeoAntSites$SSBS)) #number of sites
```

# ANTS
Here I am going to do the same thing I did with dung beetle sites, but with the ant sites
### check what interactions and levels I need to remove
```{r}
table(NeoAntSites$Predominant_land_use, NeoAntSites$Use_intensity)
```

### Based on the Ant sites table, there are a number of alterations I probably need to make before the analyses.
* Remove "Urban" as there is only 1 site
* Collapse the "Cannot Decide" and "Indeterminate Secondary Vegetation"
* Collapse land use categories together
  + Combine "Primary forest" and "Primary non-forest"
* Collapse multiple intensities together due to lack of data
  + Create a "higher use (H)" by combining Light and Intense together for all land uses
  
In this chunk I will collapse primary vegetation together and remove all NAs and Indeterminates and remove urban sites.
```{r}
NeoAntSites <- NeoAntSites %>%
  
  mutate(

     # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                  "Primary forest" = "Primary vegetation", 
                                  "Primary non-forest" = "Primary vegetation"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
)

NeoAntSites <- droplevels(subset(NeoAntSites, Predominant_land_use != "Urban"))

# take another look at the LandUse/Use intensity split
table(NeoAntSites$Predominant_land_use, NeoAntSites$Use_intensity) 
```

### Collapse all land uses so there are (ideally) two intensities, "minimal" and "higher". If not, make sure each category has at least 30 observations in it
Since I already ran some code with "LightIntense" I am going to keep it like that until I get further down.
```{r}
NeoAntSites <- NeoAntSites %>%
  mutate(LUI = interaction(Predominant_land_use, Use_intensity, sep = "_"),
         
         # collapse Primary vegetation into one category
         LUI = recode_factor(LUI, "Primary vegetation_Minimal use" = "Primary vegetation", "Primary vegetation_Light use" = "Primary vegetation", "Primary vegetation_Intense use" = "Primary vegetation"),
         
         #collapse young secondary together
         LUI = recode_factor(LUI, "Young secondary vegetation_Intense use" = "Young secondary vegetation_LightIntense", 
                             "Young secondary vegetation_Light use" = "Young secondary vegetation_LightIntense"),
         
         #collapse Cropland Itogether"
         LUI = recode_factor(LUI, "Cropland_Intense use" = "Cropland",
                             "Cropland_Light use" = "Cropland",
                             "Cropland_Minimal use" = "Cropland"),
         
         #collapse plantation forest intense and light together
         LUI = recode_factor(LUI, "Plantation forest_Intense use" = "Plantation forest_LightIntense",
                             "Plantation forest_Light use" = "Plantation forest_LightIntense"),
         
         #collapse Intermediate secondary vegetation light and intense together
         LUI = recode_factor(LUI, "Intermediate secondary vegetation_Intense use" = "Intermediate secondary vegetation_LightIntense", "Intermediate secondary vegetation_Light use" = "Intermediate secondary vegetation_LightIntense"),
         
         #collapse Pasture into light and intense
         LUI = recode_factor(LUI, "Pasture_Light use" = "Pasture",
                             "Pasture_Intense use" = "Pasture",
                             "Pasture_Minimal use" = "Pasture"),
        
  )

#remove the categories with no data
NeoAntSites <- droplevels(subset(NeoAntSites, LUI != "Mature secondary vegetation_Light use" | LUI != "Mature secondary vegetation_Intense use"))

#check to see it has worked
table(NeoAntSites$LUI)
```

### Recode the names of the LUI categories to make them shorter
Here, I am making the "LightIntense" category "Higher (H)"
```{r}
#LightIntense use category will be classified as "Higher (H)" use for brevity
NeoAntSites<- NeoAntSites%>%
  mutate(LUI = plyr::revalue(LUI, c("Primary vegetation" = "PV",
                                    "Mature secondary vegetation_Minimal use" = "MSV", #since there is only this category now, no other intensities
                                    "Intermediate secondary vegetation_Minimal use" = "ISVMU",
                                    "Intermediate secondary vegetation_LightIntense" = "ISVHU",
                                    "Young secondary vegetation_Minimal use" = "YSVMU",
                                    "Young secondary vegetation_LightIntense" = "YSVHU",
                                    "Plantation forest_Minimal use" = "PFMU",
                                    "Plantation forest_LightIntense" = "PFHU",
                                    "Pasture" = "P",
                                    "Cropland" = "C"
                                    )))
```

### reorder the factor levels
```{r}
#first get a list of the levels
levels(NeoAntSites$LUI)

#recode the levels
NeoAntSites$LUI <- factor(NeoAntSites$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

#check it worked
levels(NeoAntSites$LUI)
```

### Check to see how many sites we have in each category. There are at least 30, so the data spread should be ok
```{r}
table(NeoAntSites$LUI)
```

### Create the model data frames
```{r}
#ant model data, abundance 
model_data_ant_ab <- drop_na(NeoAntSites,
                             RescaledAbundance, Predominant_land_use, Use_intensity, LUI)

#ant log and sqrt abundances
model_data_ant_ab <- mutate(model_data_ant_ab, 
                        logAbundance = log(RescaledAbundance + 1),
                        sqrtAbundance = sqrt(RescaledAbundance))

## ant species richness data
model_data_ant_sr <- drop_na(NeoAntSites,
                             SpeciesRichness, Predominant_land_use, Use_intensity, LUI)
```

### Check the number of sources, studies, and sites in the model data
```{r}
#ants
length(unique(model_data_ant_sr$Source_ID)) #number of sources
length(unique(model_data_ant_sr$SS)) #number of studies
length(unique(model_data_ant_sr$SSBS)) #number of sites
```

### Reorder the factor levels for the model ant datasets
Reorder the levels so that visualisation is more clear (not always necessary but in case the model data gets messed up)
```{r}
### ANTS

#abundance

    #reorder abundance LUI levels
    model_data_ant_ab$LUI <- factor(model_data_ant_ab$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

    #reorder abundance Predominant land use levels
    model_data_ant_ab$Predominant_land_use <- factor(model_data_ant_ab$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))

    #reorder abundance Use intensity levels
    model_data_ant_ab$Use_intensity <- factor(model_data_ant_ab$Use_intensity, levels = c("Minimal use", "Light use", "Intense use"))

#species richness

    #reorder Species richness LUI levels
    model_data_ant_sr$LUI <- factor(model_data_ant_sr$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

    #reorder Species richness Predominant land use levels
    model_data_ant_sr$Predominant_land_use <- factor(model_data_ant_sr$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))

    #reorder Species richness Use intensity levels
    model_data_ant_sr$Use_intensity <- factor(model_data_ant_sr$Use_intensity, levels = c("Minimal use", "Light use", "Intense use"))
```

The ant abundance dataframe is a bit shorter than the SR dataframe. Check the data spread in the abundance one to see if it is even
### Check the abundance dataframe data spread
```{r}
## Check LUI spread
table(model_data_ant_ab$LUI)

## Check land use spread
table(model_data_ant_ab$Predominant_land_use)
```

Unfortunately the abundance data is pretty sparse in some areas, but it makes sense because a lot of ant measurements tend to be occurrence measurements rather than abundance ones. Because of the imbalance here, I don't think I will be able to use MSV unless I collapsed the vegetation together.

The SR dataframe is the same length as the ants one, so the data spread should be even there.

# MODELLING

### Create models of Species richness 

### Create the maximal model and proceed with backwards stepwise selection
```{r}
## Create the maximal model
m1s <- glmer(SpeciesRichness ~ LUI + (1|SS) + (1|SSB), 
            data = model_data_ant_sr,
            family = "poisson")

## Remove block as random effect
m2s <- glmer(SpeciesRichness ~ LUI + (1|SS), 
            data = model_data_ant_sr,
            family = "poisson")

## Remove study as random effect
m3s <- glmer(SpeciesRichness ~ LUI + (1|SSB), 
            data = model_data_ant_sr,
            family = "poisson",
            control = glmerControl(optimizer = "bobyqa"))

## Compare the random effects structure of the model
AIC(m1s, m2s, m3s)
anova(m1s, m2s, m3s)

## Model m1s is the best random effect structure, so keep that.

## Compare the first model to an intercept only model of Species richness
m4s <- glmer(SpeciesRichness ~ (1|SS) + (1|SSB), 
            data = model_data_ant_sr,
            family = "poisson")

## Check which model fits better
anova(m1s, m4s)
model.sel(m1s, m4s)

## It seems that model m1s is the best model.
summary(m1s)
```

### Validate the model, check diagnostics of model m1s
```{r}
## run the model function
model_plot(m1s)

## Simulate the residuals of the model using DHARMa
m1s_residuals <- simulateResiduals(fittedModel = m4s, plot = F)

## Plot them
plot(m1s_residuals)

## Test the outliers and test for dispersion
testOutliers(m1s_residuals)
testDispersion(m1s_residuals)

## Test for overdispersion
overdisp_fun(m1s)
```

Model m1s diagnostic plots look ok, and there is no overdispersion
### Check the summary statistics and significance of terms
```{r}
summary(m1s)
Anova(m1s)
```

This seems to be the best model, and I might want to make alterations to it though.

Now I want to extract the effect estimates of the model and plot them so I can visualize how the LUI is affecting ant species richness.

### Simulate the effects of the ant SR model to generate figures
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
ant_sr_eff <- FEsim(m1s, 1000)

# make the term column a factor so it can be recoded
ant_sr_eff$term <- as.factor(ant_sr_eff$term)

# recode the factors to make them shorter (so I can visualise easier)
ant_sr_eff$term <- recode_factor(ant_sr_eff$term, "(Intercept)" = "PV", 
                              "LUIMSV" = "MSV", 
                              "LUIISVMU" = "ISVMU", 
                              "LUIISVHU" = "ISVHU", 
                              "LUIYSVMU" = "YSVMU", 
                              "LUIYSVHU" = "YSVHU", 
                              "LUIPFMU" = "PFMU", 
                              "LUIPFHU" = "PFHU", 
                              "LUIP" = "P", 
                              "LUIC" = "C")

# Add in upper and lower confidence intervals 
ant_sr_eff <- ant_sr_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Generate percentage differences from the baseline (first term, Primary Vegetation (PV))
ant_sr_eff <- ant_sr_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
ant_sr_eff[1,7] <- 0 #shift median
ant_sr_eff[1,8] <- 0 #shift upper
ant_sr_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(ant_sr_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_data_ant_sr_plot <- left_join(model_data_ant_sr, ant_sr_eff)
```

### Plot the effect estimates 
```{r}
## Plot the effect estimate as the percent species richness difference from primary vegetation
ant_species_richness <- model_data_ant_sr_plot %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper)+
  geom_hline(yintercept = 0, size = 0.75, color = ("black"), linetype = 1 )+
  geom_point(size = 2)+
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), size = 0.75)+
  theme_classic()+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        legend.title = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank(),
        panel.border = element_rect(colour = "black",  fill=NA),
        axis.title.y = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))+
  xlab("Land use intensity class")+
  ylab("Species richness difference (%)")+
  theme(legend.position = "none") 
ant_species_richness
```
This should be the graph that I use for the fine scale ant species richness. In terms of writing it up:

Figure X: Ant species richness response to land use intensity (LUI) in the Neotropics
Error bars show the 95% confidence interval around each estimate. Error bars that do not cross zero are statistically significantly different from primary vegetation. Colors represent the following land use types: black (primary vegetation, PV), orange (mature secondary vegetation, MSV), light blue (intermediate secondary vegetation, ISV), green (young secondary vegetation, YSV), yellow (plantation forest, PF), dark blue (pasture), dark orange (cropland). N = 948 observations


### ant effects dataframe
```{r}
write.csv(ant_sr_eff, "Ant_species_richness_effects.csv")
```

### Export the graph
```{r}
ant_species_richness_plot <- ggarrange(ant_species_richness, ant_species_richness,
          ncol = 1, nrow = 2)
ant_species_richness_plot

ggexport(ant_species_richness_plot, "ant_broad_SR_plot.pdf")
```