---
title: "4a_Ant_Three_Guild_Model"
author: "Alex Dhond"
date: "7/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script I will load in the Excel file I made that classifies ants into 3 trophic guilds (Predator/Omnivore/Herbivore) and calculate the Species richness of the guilds

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation

## Modelling ## 
library(lme4) # creating mixed effects models
library(car) # for LRT / anova comparisons
library(merTools) # for extracting model estimates
library(MuMIn) # for assessing model fit
library(optimx) #optimizing models
library(DHARMa) #for model validation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in any required functions
```{r}
## Load in the source function for checking GVIFs - for collinearity
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}

## Load in an overdispersion function from this website to help test if the models are overdispersed.
# https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#testing-for-overdispersioncomputing-overdispersion-factor
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

### Read in the excel file and tidy it up
Here I will read in the excel file that contains a list of the species and genus of ants along with their guilds
```{r}
#read in file
ant_guilds <- read.csv("Ant_species.csv")

#convert NAs to 0s
ant_guilds[is.na(ant_guilds)] <- 0

#remove column 1 and the specific classification, leaving only the 3 tier classification
ant_guilds <- ant_guilds[, -c(1,3:12)]
```

### Create a new dataframe in which Guild becomes its own column
```{r}
## pivot longer new dataframe
ant_guilds_long <- ant_guilds %>% 
  pivot_longer(!GenusSpecies, names_to = "Guild", values_to = "count")

## get rid of rows where it is 0
ant_guilds_long <- droplevels(subset(ant_guilds_long, count !=0))

## get rid of count
ant_guilds_long <- ant_guilds_long[, -3]

## I should now be able to left join this to the Ants dataset by GenusSpecies
```

# Join the ant guilds into the larger ant dataset

### Read in the ants data set, make a new column GenusSpecies
```{r}
#read in the data
NeoAnts <- read.csv("NeoAnts.csv")

# remove rows where there is no genus information (because I cannot classify into functional/trophic guilds if there is no genus information)
NeoAnts <- droplevels(subset(NeoAnts, Genus != ""))
  
# create GenusSpecies column
NeoAnts <- NeoAnts %>%
  mutate(GenusSpecies = ifelse(Best_guess_binomial == "", paste(Genus), paste(Best_guess_binomial)))
```

### Left join the two datasets
```{r}
## join the set where guild is only a column
Ants_long <- left_join(NeoAnts, ant_guilds_long)
```

### Create the sites where guild is also kept separately
```{r}
NeoAntSites_Guild1 <- Ants_long %>%

  # pull out only the merged diversity data
  distinct(merge_ID, .keep_all = TRUE) %>%
  
  # re-make SSB and SSBS values since we've now dropped a bunch of values
  mutate(SS = paste(Source_ID, Study_number),
         SSB = paste(SS, Block),
         SSBS = paste(SSB, Site_number)) %>%
  
  # group by SSBS (each unique value corresponds to a unique site)
  group_by(SSBS, Guild) %>%
  
  # now add up all the abundance measurements within each site
  mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                 sum(merged_diversity),
                                 # if the diversity metric type isn't Abundance, then leave the TotalAbundance measurement as NA
                                 NA),
         
         # if the metric is already species richness
         SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                  # just use the value as given
                                  merged_diversity,
                                  # for abundance and occurrence measurements, count the number of unique species names that are present at the site 
                                  n_distinct(Taxon_name_entered[merged_diversity > 0])),
         
         # calculate Chao's Species Richness Estimator
         # if the diversity metric is suitable for this calculation
         ChaoRichness = ifelse(Diversity_metric_is_suitable_for_Chao == TRUE,
                               # calculate the Chao estimator
                               sum(merged_diversity > 0) + (((sum(merged_diversity == 1) * (sum(merged_diversity == 1)-1)) / (2*(sum(merged_diversity == 2)+1)))),
                               # otherwise give Chao NA
                               NA)
         ) %>%
  
  # ungroup
  ungroup() %>%
  
    # now group by Study ID
  group_by(SS, Guild) %>%
  
  # pull out some useful study-level numbers
  # maximum abundance for each study
  mutate(MaxAbundance = max(TotalAbundance),
         # number of species in the study
         SpeciesInStudy = n_distinct(Taxon_name_entered[merged_diversity > 0]),
         # assess whether the study is suitable for rarefaction
         SuitableForRarefaction = ifelse(
           # if all diversity measurements are integers
           # i.e. if you round down, it should be equal to the original number
           # and the diversity metric is suitable for Chao
           all(floor(merged_diversity) == merged_diversity) &
             Diversity_metric_is_suitable_for_Chao == TRUE,
           # then class the study as suitable for rarefaction
           TRUE,
           # otherwise it can't be used
           FALSE
           )
         ) %>%
  
  # ungroup
  ungroup()  %>%
  
  # now rescale total abundance, so that within each study, abundance varies from 0 to 1.
  mutate(RescaledAbundance = TotalAbundance/MaxAbundance,
         # for statistical modelling, we'll also calculate the square root of species abundance, although we might want to use log(x+1) transformation instead
         sqrtRescaledAbundance = sqrt(RescaledAbundance),
         logRescaledAbundance = log(RescaledAbundance + 1)
         )

# pull out unique sites
NeoAntSites_Guild1 <- NeoAntSites_Guild1 %>%
  distinct(SSBS, Guild, .keep_all = TRUE)
```

### collapse land use categories together (these should be the same categories as script 4 - ant site diversity)
```{r}
NeoAntSites_Guild1 <- NeoAntSites_Guild1 %>%
  
  mutate(

     # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                  "Primary forest" = "Primary vegetation", 
                                  "Primary non-forest" = "Primary vegetation"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
)

NeoAntSites_Guild1 <- droplevels(subset(NeoAntSites_Guild1, Predominant_land_use != "Urban"))

# take another look at the LandUse/Use intensity split
table(NeoAntSites_Guild1$Predominant_land_use, NeoAntSites_Guild1$Use_intensity) 
```

### Create LUI as a new factor that combines Land use and use intensity
```{r}
NeoAntSites_Guild1 <- NeoAntSites_Guild1 %>%
  mutate(LUI = interaction(Predominant_land_use, Use_intensity, sep = "_"),
         
         # collapse Primary vegetation into one category
         LUI = recode_factor(LUI, "Primary vegetation_Minimal use" = "Primary vegetation", "Primary vegetation_Light use" = "Primary vegetation", "Primary vegetation_Intense use" = "Primary vegetation"),
         
         #collapse young secondary together
         LUI = recode_factor(LUI, "Young secondary vegetation_Intense use" = "Young secondary vegetation_LightIntense", 
                             "Young secondary vegetation_Light use" = "Young secondary vegetation_LightIntense"),
         
         #collapse Cropland Itogether"
         LUI = recode_factor(LUI, "Cropland_Intense use" = "Cropland",
                             "Cropland_Light use" = "Cropland",
                             "Cropland_Minimal use" = "Cropland"),
         
         #collapse plantation forest intense and light together
         LUI = recode_factor(LUI, "Plantation forest_Intense use" = "Plantation forest_LightIntense",
                             "Plantation forest_Light use" = "Plantation forest_LightIntense"),
         
         #collapse Intermediate secondary vegetation light and intense together
         LUI = recode_factor(LUI, "Intermediate secondary vegetation_Intense use" = "Intermediate secondary vegetation_LightIntense", "Intermediate secondary vegetation_Light use" = "Intermediate secondary vegetation_LightIntense"),
         
         #collapse Pasture into light and intense
         LUI = recode_factor(LUI, "Pasture_Light use" = "Pasture",
                             "Pasture_Intense use" = "Pasture",
                             "Pasture_Minimal use" = "Pasture"),
        
  )

#remove the categories with no data
NeoAntSites_Guild1 <- droplevels(subset(NeoAntSites_Guild1, LUI != "Mature secondary vegetation_Light use" | LUI != "Mature secondary vegetation_Intense use"))

#check to see it has worked
table(NeoAntSites_Guild1$LUI)
```

### Recode the names of the LUI categories to make them shorter
Here, I am making the "LightIntense" category "Higher (H)"
```{r}
#LightIntense use category will be classified as "Higher (H)" use for brevity
NeoAntSites_Guild1<- NeoAntSites_Guild1%>%
  mutate(LUI = plyr::revalue(LUI, c("Primary vegetation" = "PV",
                                    "Mature secondary vegetation_Minimal use" = "MSV", #since there is only this category now, no other intensities
                                    "Intermediate secondary vegetation_Minimal use" = "ISVMU",
                                    "Intermediate secondary vegetation_LightIntense" = "ISVHU",
                                    "Young secondary vegetation_Minimal use" = "YSVMU",
                                    "Young secondary vegetation_LightIntense" = "YSVHU",
                                    "Plantation forest_Minimal use" = "PFMU",
                                    "Plantation forest_LightIntense" = "PFHU",
                                    "Pasture" = "P",
                                    "Cropland" = "C"
                                    )))
```

### reorder the factor levels
```{r}
#first get a list of the levels
levels(NeoAntSites_Guild1$LUI)

#recode the levels
NeoAntSites_Guild1$LUI <- factor(NeoAntSites_Guild1$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

#check it worked
levels(NeoAntSites_Guild1$LUI)
```

### Create a model dataframe for the sites
```{r}
model_ant_ab <- drop_na(NeoAntSites_Guild1,
                             RescaledAbundance, Predominant_land_use, Use_intensity, LUI, Guild)
#ant log and sqrt abundances
model_ant_ab <- mutate(model_ant_ab, 
                        logAbundance = log(RescaledAbundance + 1),
                        sqrtAbundance = sqrt(RescaledAbundance))
  
model_ant_sr <- drop_na(NeoAntSites_Guild1,
                        SpeciesRichness, Predominant_land_use, Use_intensity, LUI, Guild)
```

### Reorder the factor levels for the model ant datasets
Reorder the levels so that visualisation is more clear
```{r}
### ANTS

#abundance

    #reorder abundance LUI levels
    model_ant_ab$LUI <- factor(model_ant_ab$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

    #reorder abundance Predominant land use levels
    model_ant_ab$Predominant_land_use <- factor(model_ant_ab$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))

    #reorder abundance Use intensity levels
    model_ant_ab$Use_intensity <- factor(model_ant_ab$Use_intensity, levels = c("Minimal use", "Light use", "Intense use"))

#species richness

    #reorder Species richness LUI levels
    model_ant_sr$LUI <- factor(model_ant_sr$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

    #reorder Species richness Predominant land use levels
    model_ant_sr$Predominant_land_use <- factor(model_ant_sr$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))

    #reorder Species richness Use intensity levels
    model_ant_sr$Use_intensity <- factor(model_ant_sr$Use_intensity, levels = c("Minimal use", "Light use", "Intense use"))
```


### Look at the data spread
The data spread looks alright
```{r}
table(model_ant_sr$LUI)
table(model_ant_sr$Predominant_land_use, model_ant_sr$Use_intensity)
table(model_ant_sr$LUI, model_ant_sr$Guild)
```

### Clean up the guilds 
```{r}
## check structure of the data
str(model_ant_sr$Guild)

## make it a factor
model_ant_sr$Guild <- as.factor(model_ant_sr$Guild)

## look at the data spread
table(model_ant_sr$Guild, model_ant_sr$Predominant_land_use)
```

Here, the species richness column is telling me how many species of that guild are present at that site.
So, I am looking at the species richness of each guild at each site.
# SPECIES RICHNESS

### First check for collinearity between the predictor variables
I don't expect there to be any collinearity between LUI and guild, but I will check anyway
```{r}
## Test for collinearity between LUI and guild
corvif(NeoAntSites_Guild1[ , c("LUI", "Guild")])

## Test for collinearity between predominant land use and guild
corvif(NeoAntSites_Guild1[ , c("Predominant_land_use", "Guild")])
```

### Create models to test the best random effects structure
### Create the maximal model and proceed with backwards stepwise selection
```{r}
## Create the maximal model. optimizer used due to convergence warnings
m1s <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Remove block as random effect. optimizer used due to convergence warnings
m2s <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Remove study as random effect. optimizer used due to convergence warnings
m3s <- glmer(SpeciesRichness ~ LUI*Guild + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Compare the random effects structure of the model
AIC(m1s, m2s, m3s)
anova(m1s, m2s, m3s)

## Model m1s is the best random effect structure, so keep that.

## Now compare fixed effect structure - additive fixed effects
m4s <- glmer(SpeciesRichness ~ LUI + Guild + (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## remove guild as an effect
m5s <- glmer(SpeciesRichness ~ LUI + (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## remove LUI as an effect
m6s <- glmer(SpeciesRichness ~ Guild + (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Compare the first model to an intercept only model of Species richness
m7s <- glmer(SpeciesRichness ~ (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson")

## Check which model fits better
anova(m1s, m4s, m5s, m6s, m7s)
model.sel(m1s, m4s, m5s, m6s, m7s)

## It seems that model m1s is the best model.
summary(m1s)
```

### Validate the model, check diagnostics of model m1s
```{r}
## run the model function
model_plot(m1s)

## Simulate the residuals of the model using DHARMa
m1s_residuals <- simulateResiduals(fittedModel = m4s, plot = F)

## Plot them
plot(m1s_residuals)

## Test the outliers and test for dispersion
testOutliers(m1s_residuals)
testDispersion(m1s_residuals)

## Test for overdispersion
overdisp_fun(m1s)
```

The model is slightly overdispersed, so I will add in a site level random effect to help account for dispersion

### Add in another random effect to account for over dispersion
```{r}
## In case of overdispersion, add in a row-level random effect
model_ant_sr$Row_random_effect <- 1:nrow(model_ant_sr)

## Add each row (essentially site) as a random effect
m5s <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## check different optimizer
m5s_test <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl(optimizer = "bobyqa"))

## Check for overdispersion
overdisp_fun(m5s)

anova(m5s, m5s_test)
summary(m5s)
summary(m5s_test)
Anova(m5s)
```

### Check criticism plots again
```{r}
## Check the residuals using the model_plot function
model_plot(m5s)

## Simulate the residuals of the model using DHARMa
m5s_residuals <- simulateResiduals(fittedModel = m5s, plot = F)

## Plot them
plot(m5s_residuals)

## Test the outliers and test for dispersion
testOutliers(m5s_residuals)
testDispersion(m5s_residuals)
```
This model is no longer overdispersed, so it should be ok

I will now make some models for SR vs land-use to see if land-use or LUI is a better fit for the model

### Create models for SR vs Land use
### Create the maximal model and proceed with backwards stepwise selection
```{r}
## Create the maximal model. optimizer used due to convergence warnings
m6s <- glmer(SpeciesRichness ~ Predominant_land_use*Guild + (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Remove block as random effect. optimizer used due to convergence warnings
m7s <- glmer(SpeciesRichness ~ Predominant_land_use*Guild + (1|SS), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Remove study as random effect. optimizer used due to convergence warnings
m8s <- glmer(SpeciesRichness ~ Predominant_land_use*Guild + (1|SSB), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Compare the random effects structure of the model
AIC(m6s, m7s, m8s)
anova(m6s, m7s, m8s)

## Model m1s is the best random effect structure, so keep that.

## Compare the first model to an intercept only model of Species richness
m9s <- glmer(SpeciesRichness ~ (1|SS) + (1|SSB), 
            data = model_ant_sr,
            family = "poisson")

## Check which model fits better
anova(m6s, m9s)
model.sel(m6s, m9s)

## It seems that model m6s is the best model.
summary(m6s)
```

### Check diagnostics for model m6s for SR vs land use
```{r}
## Check the residuals using the model_plot function
model_plot(m6s)

## Simulate the residuals of the model using DHARMa
m6s_residuals <- simulateResiduals(fittedModel = m6s, plot = F)

## Plot them
plot(m6s_residuals)

## Test the outliers and test for dispersion
testOutliers(m6s_residuals)
testDispersion(m6s_residuals)

## test for overdispersion
overdisp_fun(m6s)
```

### Once again the model is overdispersed, so add in a site level random effect to account for overdispersion
```{r}
## Add each row (essentially site) as a random effect
m10s <- glmer(SpeciesRichness ~ Predominant_land_use*Guild + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Check for overdispersion
overdisp_fun(m10s)
```

### Check criticism plots again
```{r}
## Check the residuals using the model_plot function
model_plot(m10s)

## Simulate the residuals of the model using DHARMa
m10s_residuals <- simulateResiduals(fittedModel = m10s, plot = F)

## Plot them
plot(m10s_residuals)

## Test the outliers and test for dispersion
testOutliers(m10s_residuals)
testDispersion(m10s_residuals)
```

For the model of LUI vs species richness, model m6s is the most significant. For the model of land use vs species richness, model m10s is the best

### Compare the land use model (m5s) with the LUI model (m10s) for species richness
```{r}
## Compare models
anova(m5s, m10s)
model.sel(m5s, m10s)
Anova(m5s)
## The output of both of the tests says that model m5s (the LUI model) fits the data better, which does make sense as it is the more specific model.

anova(m5s, m4s)
Anova(m5s)
anova(m5s)
## the interaction between LUI and guild explained a significant amount of variation. chi squared = 220.57, df = 19, p < 2.2e-16
```

So, that means model m9s is the model that I am going to extract the estimates from and plot. I need to figure out how to extract the estimates for interaction terms and plot them on the same graph.

### Create multiple models with different reference categories
```{r}
## ## relevel herbivores and model herbivores
model_ant_sr <- within(model_ant_sr, Guild <- relevel(Guild, ref = "Pseudo...herbivorous"))
m5s_herb <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS) + (1|SSB) + (1|Row_random_effect), 
            data = model_ant_sr,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Create omnivore dataframe and model
model_ant_sr_omni <- within(model_ant_sr, Guild <- relevel(Guild, ref = "Omnivore"))
m5s_omni <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS) + (1|SSB) + (1|Row_random_effect), data = model_ant_sr_omni,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## Create predator dataframe and model
model_ant_sr_pred <- within(model_ant_sr, Guild <- relevel(Guild, ref = "Predator"))
m5s_pred <- glmer(SpeciesRichness ~ LUI*Guild + (1|SS) + (1|SSB) + (1|Row_random_effect), data = model_ant_sr_pred,
            family = "poisson",
            control = glmerControl("optimx", optCtrl=list(method="nlminb")))

## check model summaries
summary(m5s_herb)
summary(m5s_omni)
summary(m5s_pred)
```

### Simulate the effects of the trophic guild SR model to generate figures for omnivores
```{r}
# gather the effects and confidence intervals using simulation. Simulate 10000 times
omni_eff <- FEsim(m5s_omni, 1000)

# make the term column a factor so it can be recoded
omni_eff$term <- as.factor(omni_eff$term)

# take only omnivores
omni_eff <- omni_eff %>%
  slice(1:10)

# recode the factors to make them shorter (so I can visualise easier)
omni_eff$term <- recode_factor(omni_eff$term, "(Intercept)" = "PV", 
                              "LUIMSV" = "MSV", 
                              "LUIISVMU" = "ISVMU", 
                              "LUIISVHU" = "ISVHU", 
                              "LUIYSVMU" = "YSVMU", 
                              "LUIYSVHU" = "YSVHU", 
                              "LUIPFMU" = "PFMU", 
                              "LUIPFHU" = "PFHU", 
                              "LUIP" = "P", 
                              "LUIC" = "C")

# Add in upper and lower confidence intervals 
omni_eff <- omni_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Generate percentage differences from the baseline (first term, Primary Vegetation (PV))
omni_eff <- omni_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
omni_eff[1,7] <- 0 #shift median
omni_eff[1,8] <- 0 #shift upper
omni_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(omni_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_ant_omni_plot <- left_join(model_ant_sr, omni_eff)
```

### Create the Predator plot data
```{r}
# gather the effects and confidence intervals using simulation. Simulate 10000 times
pred_eff <- FEsim(m5s_pred, 1000)

# make the term column a factor so it can be recoded
pred_eff$term <- as.factor(pred_eff$term)

# take only omnivores
pred_eff <- pred_eff %>%
  slice(1:10)

# recode the factors to make them shorter (so I can visualise easier)
pred_eff$term <- recode_factor(pred_eff$term, "(Intercept)" = "PV", 
                              "LUIMSV" = "MSV", 
                              "LUIISVMU" = "ISVMU", 
                              "LUIISVHU" = "ISVHU", 
                              "LUIYSVMU" = "YSVMU", 
                              "LUIYSVHU" = "YSVHU", 
                              "LUIPFMU" = "PFMU", 
                              "LUIPFHU" = "PFHU", 
                              "LUIP" = "P", 
                              "LUIC" = "C")

# Add in upper and lower confidence intervals 
pred_eff <- pred_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Generate percentage differences from the baseline (first term, Primary Vegetation (PV))
pred_eff <- pred_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
pred_eff[1,7] <- 0 #shift median
pred_eff[1,8] <- 0 #shift upper
pred_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(pred_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_ant_pred_plot <- left_join(model_ant_sr, pred_eff)
```

### Create the Herbivore plot data
```{r}
# gather the effects and confidence intervals using simulation. Simulate 10000 times
herb_eff <- FEsim(m5s_herb, 1000)

# make the term column a factor so it can be recoded
herb_eff$term <- as.factor(herb_eff$term)

herb_eff <- herb_eff %>%
  slice(1:10)

# recode the factors to make them shorter (so I can visualise easier)
herb_eff$term <- recode_factor(herb_eff$term, "(Intercept)" = "PV", 
                              "LUIMSV" = "MSV", 
                              "LUIISVMU" = "ISVMU", 
                              "LUIISVHU" = "ISVHU", 
                              "LUIYSVMU" = "YSVMU", 
                              "LUIYSVHU" = "YSVHU", 
                              "LUIPFMU" = "PFMU", 
                              "LUIPFHU" = "PFHU", 
                              "LUIP" = "P", 
                              "LUIC" = "C")

# Add in upper and lower confidence intervals 
herb_eff <- herb_eff %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Generate percentage differences from the baseline (first term, Primary Vegetation (PV))
herb_eff <- herb_eff %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci)) / exp(median[1]))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci)) / exp(median[1]))*100)-100))

# Shift the baseline down to 0
herb_eff[1,7] <- 0 #shift median
herb_eff[1,8] <- 0 #shift upper
herb_eff[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(herb_eff)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
model_ant_herb_plot <- left_join(model_ant_sr, herb_eff)
```

Regardless, combine them all into one plot

### Combine all estimates into one
```{r}
## Join them
ant_trophic_SR_eff <- rbind(herb_eff, omni_eff, pred_eff)

## Make sure the guilds are correct
ant_trophic_SR_eff$Guild <- c("Herbivore","Herbivore","Herbivore","Herbivore","Herbivore","Herbivore","Herbivore","Herbivore","Herbivore","Herbivore","Omnivore","Omnivore","Omnivore","Omnivore","Omnivore","Omnivore","Omnivore","Omnivore","Omnivore","Omnivore","Predator","Predator","Predator","Predator","Predator","Predator","Predator","Predator","Predator","Predator")
```

### Plot all of the effects together
```{r}
## set the limits 
ant_3_X_Limits <- c("PV",
              "MSV",
              "ISVMU",
              "ISVHU",
              "YSVMU",
              "YSVHU",
              "PFMU",
              "PFHU",
              "P",
              "C")

## create color scheme
guild_3_tier_colors <- c("green4", "dodgerblue", "indianred2")

## Plot the effect estimate as the percent species richness difference from primary vegetation
guild_3_tier_plot <- ant_trophic_SR_eff %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper, colour = Guild)+
  geom_hline(yintercept = 0, size = 0.5, color = ("black"), linetype = 1 )+
  geom_point(size = 2, position = position_dodge(width = 0.75))+
  geom_vline(xintercept = 1.5, linetype = 2)+
  geom_vline(xintercept = 2.5, linetype = 2)+
  geom_vline(xintercept = 3.5, linetype = 2)+
  geom_vline(xintercept = 4.5, linetype = 2)+
  geom_vline(xintercept = 5.5, linetype = 2)+
  geom_vline(xintercept = 6.5, linetype = 2)+
  geom_vline(xintercept = 7.5, linetype = 2)+
  geom_vline(xintercept = 8.5, linetype = 2)+
  geom_vline(xintercept = 9.5, linetype = 2)+
  scale_x_discrete(limits=ant_3_X_Limits, labels=c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))+
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width=0.75), size = 0.75)+
  theme_classic()+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        panel.border = element_rect(colour = "black",  fill=NA),
        axis.title.y = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))+
  xlab("Land use intensity class")+
  ylab("Species richness difference (%)")+
  scale_colour_manual(values=guild_3_tier_colors)
guild_3_tier_plot
```

### generate the legend
```{r}
## Create a separate plot, make a bar graph, export the legend only
ant_3_guild_legend <- ggplot(ant_trophic_SR_eff, aes(x = LUI, fill = Guild)) + 
  scale_fill_manual(values=guild_3_tier_colors)+
  guides(fill=guide_legend(title.position = "top", nrow=1,byrow=TRUE))+
  theme(legend.position = "bottom",
        legend.text = element_text(face = "bold", size = 8),
        legend.title = element_text(face = "bold"))+
  geom_bar()
ant_3_guild_legend

# Using the cowplot package
ant_3_legend <- cowplot::get_legend(ant_3_guild_legend)
ant_3_legend <- as_ggplot(ant_3_legend)
ant_3_legend
```

## combine the legend and the plot and export 
```{r}
Ant_SR_3_plots <- ggarrange(guild_3_tier_plot, ant_3_legend,
                    ncol = 1, nrow = 2)
Ant_SR_3_plots
ggexport(Ant_SR_3_plots, filename = "Ant_SR_3_plots.pdf")
```
