---
title: "4c_Ant_Nine_Guild_Simpsons"
author: "Alex Dhond"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for calculating Simpson's diversity of the nine trophic guilds, which gives me a measure of the effective number of trophic guilds represented at a site or land use. In this script I also model and plot the output

### For clearing the workspace prior to running
```{r}
rm(list=ls())
```

### Load the required packages
```{r}
## Data manipulation ## 
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(vegan) # calculating simpsons index

## Modelling ## 
library(lme4) # creating mixed effects models
library(car) # for LRT / anova comparisons
library(merTools) # for extracting model estimates
library(MuMIn) # for assessing model fit
library(optimx) #optimizing models
library(DHARMa) #for model validation

## Plotting ## 
library(ggplot2) # basic plotting package
library(ggpubr) # for arranging and exporting plots
library(cowplot) # for arranging and exporting plots
```

### Load in any required functions
```{r}
## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
   # set up a 2 x 2 grid for plotting
  par(mfrow = c(1,3))
  par(ask = TRUE) 
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

### Read in the 9 tier ant simpson model data
```{r}
nine_trophic_tier_model_ant_sr <- read.csv("9_guild_ant_model_sr.csv")
```

### Create a new dataframe for Simpson's diversity where I can extract the guilds and the sites

There are a few things to note here: 
1. I first need to get rid of all the sites where there are no species of any guild (ie zeroes across the entire matrix) as these are completely uninformative

2. Next, I need to remove all rows where there is only 1 species of each guild at a site. These singleton sites are also uninformative because I do not have any abundance data, so I do not know if there is actually complete dominance at this site (which is what it would indicate)

```{r}
## When I group by SSBS and Guild, it gives me the number of species in each guild at each site. So, at a given site, I could have X number of Predators, Y number of Omnivores, and Z number of Herbivorous ants. 

## create initial dataframe
Simpson_df <- nine_trophic_tier_model_ant_sr %>%
  dplyr::select(SSBS, SpeciesRichness, Guild)

## make it wide data so each guild is its own column
Simpson_df_9_tier <- Simpson_df %>%
  pivot_wider(names_from = Guild, values_from = SpeciesRichness)

## remove NAs
Simpson_df_9_tier[is.na(Simpson_df_9_tier)] <- 0

## make site a factor
Simpson_df_9_tier$SSBS <- as.factor(Simpson_df_9_tier$SSBS)

## Create a dataframe where all of the sites that did not contain any of the three guilds are excluded
Simp_zeros_removed <- Simpson_df_9_tier %>%
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0)

948 - 912
## Removed 36 rows where there was nothing

## Create a dataframe where all of the sites that had only 1 guild at them are also excluded.
Simp_no_ones_zeros <- Simpson_df_9_tier %>%
  
  ## where they are all 0
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## where arboreal omnivores are the only one
  filter(`Arboreal omnivores` != 1 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## where gen pred are the only 1
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 1 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## when spec pred are the only 1
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 1 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## when ground omnivores are the only 1
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 1 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## when RHP are the only 1
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=1 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## when LC are the only one 
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=1 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## where CA are the only 1
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 1 | `Arboreal predators` !=0 | `Generalist omnivores` != 0) %>%
  
  ## where Ap are the one
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=1 | `Generalist omnivores` != 0) %>%
  
  ## where GO are the one
  filter(`Arboreal omnivores` != 0 | `Ground-dwelling generalist predators` != 0 | `Ground-dwelling specialist predators` != 0 | `Ground-dwelling omnivores` != 0 | `Raid-hunting predators` !=0 | `Leaf-cutters` !=0 | `Cryptobiotic attines` != 0 | `Arboreal predators` !=0 | `Generalist omnivores` != 1)
  
  
948 - 877
## This removes 71 rows, where there was only 1 species of each guild present at the site

## I want to only look at simpsons inverse index

#generate inverse simpsons index  for larger dataset
inv_simp_all <- plyr::ddply(Simpson_df_9_tier, ~SSBS, function(x) {
  data.frame(INVSIMPSON=diversity(x[-1], index="invsimpson"))
})

## generate inv simpson for no zeros dataset
inv_simp_no_zero <- plyr::ddply(Simp_zeros_removed, ~SSBS, function(x) {
  data.frame(INVSIMPSON=diversity(x[-1], index="invsimpson"))
})

## generate inv simpson for no ones or zeros dataset
inv_simp_no_one_zero <- plyr::ddply(Simp_no_ones_zeros, ~SSBS, function(x) {
  data.frame(INVSIMPSON=diversity(x[-1], index="invsimpson"))
})

## Add in log + 1 metrics in case of non-normality
inv_simp_all$LOGINVSIMPSON <- log(inv_simp_all$INVSIMPSON + 1)
inv_simp_no_zero$LOGINVSIMPSON <- log(inv_simp_no_zero$INVSIMPSON + 1)
inv_simp_no_one_zero$LOGINVSIMPSON <- log(inv_simp_no_one_zero$INVSIMPSON + 1)
```

### Left join Simpson's back into the model data
```{r}
## join by SSBS

## no zeros data set 
model_sr_no_zero <- left_join(nine_trophic_tier_model_ant_sr, inv_simp_no_zero)

## no zeros or ones data set
model_9_tier_sr_no_one_no_zero <- left_join(nine_trophic_tier_model_ant_sr, inv_simp_no_one_zero)
```

### Clean up the no ones or zeros dataset
```{r}
## Drop all NAs from relevant modelling columns
model_9_tier_sr_no_one_no_zero <- drop_na(model_9_tier_sr_no_one_no_zero, 
                                   Predominant_land_use, LUI, INVSIMPSON, LOGINVSIMPSON)

## Relevel LUI factors 
model_9_tier_sr_no_one_no_zero$LUI <- factor(model_9_tier_sr_no_one_no_zero$LUI, levels = c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C"))

## Relevel land use factors
model_9_tier_sr_no_one_no_zero$Predominant_land_use <- factor(model_9_tier_sr_no_one_no_zero$Predominant_land_use, levels = c("Primary vegetation", "Mature secondary vegetation", "Intermediate secondary vegetation", "Young secondary vegetation", "Plantation forest", "Pasture", "Cropland"))
```


# MODEL 1 - no zeros or ones

### Model the no ones or zeros dataset with no log transformation
```{r}
## Make maximal model and do backwards stepwise selection
modsimp1 <- lmer(INVSIMPSON ~ LUI + (1|SS) + (1|SSB), data = model_9_tier_sr_no_one_no_zero)

## only study as random effect
modsimp2 <- lmer(INVSIMPSON ~ LUI + (1|SS), data = model_9_tier_sr_no_one_no_zero)

## block as random effect
modsimp3 <- lmer(INVSIMPSON ~ LUI + (1|SSB), data = model_9_tier_sr_no_one_no_zero)

## compare random effects structure
AIC(modsimp1, modsimp2, modsimp3)
model.sel(modsimp1, modsimp2, modsimp3)

## Intercept only model with study and block as random effect
modsimp4 <- lmer(INVSIMPSON ~ (1|SS) + (1|SSB), data = model_9_tier_sr_no_one_no_zero)

## compare it to null model
AIC(modsimp1, modsimp4)

## model 1 seems to be the best
```

### Check the model diagnostic plots
Model modelsimp1 is the best - look at diagnostic plots
```{r}
## run the model function
model_plot(modsimp1)

## Simulate the residuals of the model using DHARMa
modsimp1_residuals <- simulateResiduals(fittedModel = modsimp1, plot = F)

## Plot them
plot(modsimp1_residuals)

## Test the outliers and test for dispersion
testOutliers(modsimp1_residuals)
testDispersion(modsimp1_residuals)
```

It looks ok, now I will make the log transformed models as well
### Create log transformed models as well to improve normality
```{r}
## Make maximal model and do backwards stepwise selection
modsimp5 <- lmer(LOGINVSIMPSON ~ LUI + (1|SS) + (1|SSB), data = model_9_tier_sr_no_one_no_zero)

## only study as random effect
modsimp6 <- lmer(LOGINVSIMPSON ~ LUI + (1|SS), data = model_9_tier_sr_no_one_no_zero)

## block as random effect
modsimp7 <- lmer(LOGINVSIMPSON ~ LUI + (1|SSB), data = model_9_tier_sr_no_one_no_zero)

## compare random effects structure
AIC(modsimp5, modsimp6, modsimp7)
model.sel(modsimp5, modsimp6, modsimp7)

## Intercept only model with study and block as random effect
modsimp8 <- lmer(LOGINVSIMPSON ~ (1|SS) + (1|SSB), data = model_9_tier_sr_no_one_no_zero)

## compare it to null model
AIC(modsimp5, modsimp10)
AIC(modsimp9, modsimp10)
Anova(modsimp5)
## model 5 seems to be the best
```

### Check model diagnostics for model 5
```{r}
## run the model function
model_plot(modsimp5)

## Simulate the residuals of the model using DHARMa
modsimp5_residuals <- simulateResiduals(fittedModel = modsimp5, plot = F)

## Plot them
plot(modsimp5_residuals)

## Test the outliers and test for dispersion
testOutliers(modsimp5_residuals)
testDispersion(modsimp5_residuals)
```

Log transforming normalizes it a little bit, but not by much. I will keep the log transformed version

### Gather the estimates and significance of terms
```{r}
Anova(modsimp5)
summary(modsimp5)
```

### Gather the effects and plot the log transformed model
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
simp_9_eff_1 <- FEsim(modsimp5, 1000)

# make the term column a factor so it can be recoded
simp_9_eff_1$term <- as.factor(simp_9_eff_1$term)

# recode the factors to make them shorter (so I can visualise easier)
simp_9_eff_1$term <- recode_factor(simp_9_eff_1$term, "(Intercept)" = "PV", 
                              "LUIMSV" = "MSV", 
                              "LUIISVMU" = "ISVMU", 
                              "LUIISVHU" = "ISVHU", 
                              "LUIYSVMU" = "YSVMU", 
                              "LUIYSVHU" = "YSVHU", 
                              "LUIPFMU" = "PFMU", 
                              "LUIPFHU" = "PFHU", 
                              "LUIP" = "P", 
                              "LUIC" = "C")

# Add in upper and lower confidence intervals 
simp_9_eff_1 <- simp_9_eff_1 %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

# Back transform the log simpson estimates so that I get actual values
simp_9_eff_1 <- simp_9_eff_1 %>%
 mutate(
    Percent_diff = ((((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100)) %>%
  mutate(
    Percent_upper = ((((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1))*100)-100)) %>%
  mutate(
    Percent_lower = ((((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1))*100)-100))

# Shift the baseline down to 0
simp_9_eff_1[1,7] <- 0 #shift median
simp_9_eff_1[1,8] <- 0 #shift upper
simp_9_eff_1[1,9] <- 0 #shift lower

# Rename the sim eff term column name to LUI for joining
colnames(simp_9_eff_1)[1] <- "LUI"

# Left join the sim eff to the model data to make graphing easier
simp_eff_1_plot <- left_join(model_9_tier_sr_no_one_no_zero, simp_9_eff_1)
```

### Create the x-axis labels
```{r}
X_Limits <- c("PV", "MSV", "ISVMU", "ISVHU", "YSVMU", "YSVHU", "PFMU", "PFHU", "P", "C")
```

### Plot the nine guild simpsons diversity
```{r}
guild_9_tier_simpson <- simp_eff_1_plot %>%
  ggplot()+
  aes(x = LUI, y = Percent_diff, ymin = Percent_lower, ymax = Percent_upper)+
  geom_hline(yintercept = 0, size = 0.75, color = ("black"), linetype = 1 )+
  geom_point(size = 2, position = position_dodge(width = 0.5))+
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width=0.5), size = 0.75)+
  theme_classic()+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        legend.title = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank(),
        panel.border = element_rect(colour = "black",  fill=NA),
        axis.title.y = element_text(face = "bold"))+
  xlab("Land use intensity class")+
  ylab("Simpson's diversity difference (%)")
guild_9_tier_simpson
```
### Export the plot
```{r}
## Export the graph
simpson_nine_tier_plot <- ggarrange(guild_9_tier_simpson, guild_9_tier_simpson,
                    ncol = 1, nrow = 2)
ggexport(simpson_nine_tier_plot, filename = "simpson_nine_tier_plot.pdf")
```