---
title: "Ant_Species_List"
author: "Alex Dhond"
date: "7/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script I will create and export a list of all the ant genus and species so that I can classify them according to trophic classifications

# 1. For Ants:
  + Generate a list of all the ant species present at the Neotropical sites I am looking at
  + Take this list, export it as an Excel file or CSV, and then match the Genus/Species to the trophic guild categorization of Groc et al. 2014

### Read in the ant data
```{r}
NeoAnts <- read.csv("NeoAnts.csv")
```

### Remove where there is no genera information and create new column
```{r}
## drop missing genera information
NeoAnts <- droplevels(subset(NeoAnts, Genus != ""))

##  create a new column called GenusSpecies
NeoAnts <- NeoAnts %>%
  mutate(GenusSpecies = ifelse(Best_guess_binomial == "", paste(Genus), paste(Best_guess_binomial)))

## Make it a factor
NeoAnts$GenusSpecies <- as.factor(NeoAnts$GenusSpecies)

## Export just the GenusSpecies column
Ant_species <- as.factor(levels(NeoAnts$GenusSpecies))

## Save it as a file
write.csv(Ant_species, "Ant_species.csv")
```